---
title: "Effectiveness data preparation"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=F, message=F}
library(here)
library(tidyverse)
library(dplyr)
suppressPackageStartupMessages(library(tidylog))
library(ggplot2)
library(RColorBrewer)
library(ggdist)
library(plotly)
library(effect.lndscp)
library(knitr)
library(patchwork)

theme_set(theme_minimal())
```

# Seed Dispersal Effectiveness data

Load Quantity (QTC) and Quality (QLC) datasets:
```{r}
qtc.raw <- readRDS(here("data-raw/SDE_QTY_Nfruits_escape.rds"))

sde.qlc.raw <- readRDS(here("data-raw/SDE_QLTY_mh.rds"))
```

Convert QTC data to wide format; where columns represent plant individuals and each row is one entry for a bird (n=1000 per bird)
```{r}
sde.qtc.wide <- qtc.raw %>% select(bird_sp, id_plant, qty.fru) %>%
  pivot_wider(names_from = id_plant, values_from = qty.fru) %>%
  tidyr::unnest() %>% arrange(bird_sp)
```

Convert QLC data to wide format; where there is one row per species, and mh effects are in columns.

Leave just one value for the prob to escape bird predation (Prob.escape.bird.pred), as is equal in all microhabitats.

Then sum the probabilities in all mh to calculate total bird quality.
```{r}
sde.qlc.wide <- sde.qlc.raw %>% 
  select(bird_sp, mh, Prob.escape.bird.pred, Prob.bird.disperse.mh, 
         Prob.escape.rodent.pred, Prob.seedling.surv, qlty.mh) %>%
  pivot_wider(names_from = mh, values_from = c(Prob.escape.bird.pred, Prob.bird.disperse.mh, 
              Prob.escape.rodent.pred, Prob.seedling.surv, qlty.mh)) %>% 
  tidyr::unnest() %>% select(-c(Prob.escape.bird.pred_FR, Prob.escape.bird.pred_PP, 
                                Prob.escape.bird.pred_NF, Prob.escape.bird.pred_OA)) %>% 
  rename(Prob.escape.bird.pred = Prob.escape.bird.pred_PL) %>%
  mutate(sde.qlc=qlty.mh_FR+qlty.mh_PL+qlty.mh_NF+qlty.mh_OA+qlty.mh_PP) %>%
  arrange(bird_sp)
```

Check bird quality ranking (mean and median)
```{r}
sde.qlc.wide %>% group_by(bird_sp) %>% 
  summarise(median=median(sde.qlc), mean=mean(sde.qlc)) %>% 
  arrange(desc(mean))
```


Merge SDE quantity and quality in wide format:
```{r}
sde.wide <- cbind(sde.qtc.wide, sde.qlc.wide)
```

Convert SDE to long format; where each row reseprent one interaction
```{r}
sde.long <- sde.wide %>% 
  pivot_longer(cols=c(2:81), names_to="id_plant", values_to="qty.fru") %>% drop_na(qty.fru) %>%
  relocate(id_plant, .before=bird_sp) %>%
  relocate(qty.fru, .after=bird_sp) %>%
  relocate(sde.qlc, .after=qty.fru) %>% 
  arrange(id_plant, bird_sp)
```


# Resource Provisioning Effectiveness

Correct fruit consumption in seed predators by substracting the seed escaping predation (i.e. seeds fallen from the beak of predators)
```{r}
rpe.qtc <- qtc.raw %>% 
  mutate(qty.fru.consumed = ifelse(bird_sp %in% c("Chloris chloris", "Fringilla coelebs", 
           "Pyrrhula pyrrhula", "Coccothrarustes coccothrarustes"), 
           qty.fru*(1-prob.escape), qty.fru))
```

Convert quantity to long format where each row is a plant (n=1000 simulations per plant) and columns are birds
```{r}
rpe.qtc.wide <- rpe.qtc %>% select(bird_sp, id_plant, qty.fru.consumed) %>%
  pivot_wider(names_from = bird_sp, values_from = qty.fru.consumed) %>%
  tidyr::unnest() %>% arrange(id_plant)
```


Load observed pulp and seed weights:
convert data to mg and calculate pulp weight
```{r}
puntal_fruits <- read.csv(here('data-raw/fruit_characteristics.csv'), sep=";", comment.char = "#")%>% 
  mutate(id_plant = as.factor (plant_id))%>%
  mutate(fruit_weight=fruit_weight/10000) %>%
  mutate(seed_weight=seed_weight/10000) %>% 
  mutate(pulp_weight=(fruit_weight-seed_weight))
```

Generate 1000 entries for pulp and seed weights values for each plant:
```{r}
set.seed(1)

plants <- c(301:327, 329:332, 334:370, 372:376, 378:383)

# plants 328, 333, 377 (did not fruit in 2nd field season, so no interaction data), and plant 371 has no pulp data

weight_imputed <- data.frame()

for (i in 1:length(plants)){
plant_pulp <- puntal_fruits %>% filter(id_plant==plants[[i]])
id_plant = rep (plants[[i]], 1000)
pulp_weight = sample(na.omit(plant_pulp$pulp_weight), size = 1000, replace=TRUE)
seed_weight = sample(na.omit(plant_pulp$seed_weight), size = 1000, replace=TRUE)
weight_plant <- cbind(id_plant, pulp_weight, seed_weight)
weight_imputed <- rbind(weight_imputed, weight_plant)
}

#add NA info for plant 371
pulp_weight = sample(NA, size = 1000, replace=TRUE)
seed_weight = sample(NA, size = 1000, replace=TRUE)
weight_plant_371 <- cbind(id_plant=371, pulp_weight, seed_weight)
weight_imputed <- rbind(weight_imputed, weight_plant_371)
```

Compare distributions with observed data and repetitions. They maintain the same distribution.
```{r}
#histogram pulp mass distribution vs. IMPUTED pulp mass distribution
puntal_fruits %>%
  ggplot(aes(x=pulp_weight)) +
  geom_histogram(color="red", fill="white", binwidth=0.001) +
  geom_histogram(aes(x=seed_weight), color="grey", fill="white", binwidth=0.001, alpha=0.1)

weight_imputed %>%
  ggplot(aes(x=pulp_weight))  +
  geom_histogram(color="red", fill="white", binwidth=0.001) +
  geom_histogram(aes(x=seed_weight), color="grey", fill="white", binwidth=0.001, alpha=0.1)

```

Load % of water and energy values
```{r}
HR_fruits <- read.csv(here('data-raw/fruit_traits_HR_pedro.csv'), sep=";", comment.char = "#")

PWP <- mean(HR_fruits$pcap)
SWP <- mean(HR_fruits$pcas)

P.energy <- 25.2513 #Using nutrient percentages of Herrera 1987 Ecol Monog  (58.8 for lipids, 5.5. for prots and 25.8 for NSC) and using conversion factors of FAO - MacLean 2003 which are: 14.1 for prots, 15.1 for carbs and 35 for lipids. In Herrera 1982 Evol he uses different conversion factors, which are: 23.63 for prots, 17.15 for carbs and 39.53 for lipids (rendering 7.23 Kcal/g = 30.25032 kJ/g)
S.energy <- 28.14 #From Khiari t al. 2020 Biomass Conversion & Biorefinery in MJ/kg = kJ/g
```

Calculate total fruit reward (fresh weight * (1-water %) * energy/dry g)
```{r}
rpe.qlc <- weight_imputed %>% 
  mutate(PWP=PWP, SWP=SWP, P.energy=P.energy, S.energy=S.energy) %>%
  mutate(p.reward = pulp_weight*(1-PWP)*P.energy,
         s.reward = seed_weight*(1-SWP)*S.energy) %>% #kj/fruit obtained
  mutate(id_plant=as.factor(id_plant)) %>% arrange(id_plant)
```

Summary of pulp and seed reward:
```{r}
summary(rpe.qlc$p.reward)
summary(rpe.qlc$s.reward)
```

Merge RPE quantity and quality in wide format:
```{r}
rpe.wide <- cbind(rpe.qtc.wide, rpe.qlc)
```

Convert RPE to long format:
```{r}
rpe.long <- rpe.wide %>%
  pivot_longer(cols = `Chloris chloris`:`Turdus viscivorus`, 
               names_to = "bird_sp", 
               values_to = "fruits.consumed.sp") %>%
  filter(!is.na(fruits.consumed.sp)) %>% 
  arrange(id_plant, bird_sp)

missing_qtc_attributes <- rpe.qtc %>%
  arrange(id_plant, bird_sp) %>%
  select(site, Nvisit.site, Pvisit.sp, Pfeed, Nfruits, qty.fru, qty.fru.consumed)

rpe.long <- cbind(missing_qtc_attributes, rpe.long) %>%
  mutate(fruit.reward = ifelse(bird_sp %in% c("Chloris chloris", "Fringilla coelebs", "Pyrrhula pyrrhula",
                                              "Coccothraustes coccothraustes"), s.reward, p.reward)) %>% 
  relocate(id_plant, .after=site) %>%
  relocate(bird_sp, .after=id_plant) %>%
  relocate(qty.fru.consumed, .after=bird_sp) %>%
  relocate(fruit.reward, .after=qty.fru.consumed) %>%
  select(-fruits.consumed.sp) %>%
  arrange(id_plant, bird_sp)
```


# MERGE SDE AND RPE

Remove columns in SDE with the same name present in RPE data, and rename the qty fruits as "manipulated".
```{r}
sde.long.selec <- sde.long %>% select(-c(id_plant, bird_sp)) %>% rename(qty.fru.manipulated=qty.fru)
```

Merge dataset already ordered by id_plant and bird_sp
```{r}
effect.long <- cbind(rpe.long, sde.long.selec) %>%
  relocate(qty.fru.manipulated, .after=bird_sp) %>%
  relocate(sde.qlc, .after=fruit.reward) %>%
  relocate(qty.fru.manipulated, .after=bird_sp)
```

# ADD INDIVIDUAL PLANTS TRAITS
Load initcrop and areas:
```{r}
dron_areas <- read.csv(here("data-raw/dron_areas.csv"), sep=";") %>% 
  mutate(id_plant=as.factor(plant_id)) %>% 
  select(-c(site, plant_id)) 

crop.consumed.wide <- readRDS(here("data-raw/SDE_QTY_crop_consumed_wide.rds")) %>% 
  mutate(id_plant=as.factor(id_plant)) 

initcrop <- crop.consumed.wide %>% group_by(id_plant) %>% summarise(initcrop=first(initcrop))
```

Add initcrop and areas and calculate SDE and RPE
```{r}
effect.long <- effect.long %>% 
  select(-qty.fru) %>% #remove duplicated variable (=qty.fru.manipulated)
  mutate(id_plant=as.factor(id_plant)) %>%
  left_join(dron_areas) %>%
  left_join(initcrop) %>%
  mutate(sde=qty.fru.manipulated*sde.qlc,
         rpe=qty.fru.consumed*fruit.reward)
```

Save dataset:
```{r}
saveRDS(effect.long, here("data", "effectiveness.rds"))
```


# SUMMARIZED DATASET

*Summarize 1000 bayes posterior distributions*
```{r}
effect.sum <- effect.long %>%
  group_by(id_plant, bird_sp) %>%
  summarise(mean.sde.qtc=mean(qty.fru.manipulated), median.sde.qtc=median(qty.fru.manipulated),
            mean.sde.qlc=mean(sde.qlc), median.sde.qlc=median(sde.qlc),
            mean.rpe.qtc=mean(qty.fru.consumed), median.rpe.qtc=median(qty.fru.consumed),
            mean.rpe.qlc=mean(fruit.reward), median.rpe.qlc=median(fruit.reward),
            initcrop=first(initcrop), area=first(area_new),
            median.sde.qlc.FR=median(qlty.mh_FR), median.sde.qlc.NF=median(qlty.mh_NF), 
            median.sde.qlc.OA=median(qlty.mh_OA), median.sde.qlc.PP=median(qlty.mh_PP), 
            median.sde.qlc.PL=median(qlty.mh_PL),
            mean.sde=mean(sde), median.sde=median(sde),
            mean.rpe=mean(rpe), median.rpe=median(rpe))
```

Save summarized dataset:
```{r}
write_csv(effect.sum, here("data", "effectiveness_summary.csv"))
```


*Generate interaction matrix with Quantity data (i.e. number of fruits consumed)*
```{r}
effect.sum.mat <- effect.sum %>% 
  separate(bird_sp, into = c("genus", "sp"), remove = FALSE) %>% 
  mutate(bird_sp_short = paste0(str_sub(genus, end = 3), ".", str_sub(sp, end = 3))) %>% 
  mutate(bird_sp_short = ifelse(bird_sp_short == "NA.NA", NA, bird_sp_short)) %>% 
  select(-genus, -sp) %>% 
  relocate(bird_sp_short, .after = bird_sp)

int_matrix <- effect.sum.mat %>% select(bird_sp_short, id_plant, median.rpe.qtc) %>%
  mutate(median.rpe.qtc=round(median.rpe.qtc, digits = 2)) %>%
  pivot_wider(names_from = bird_sp_short, values_from = median.rpe.qtc)
```

```{r}
write_csv(int_matrix, here("data", "consumed_fruit_mat.csv"))
```


