---
title: "Effectiveness correlation distribution"
author: "Elena Quintero"
output: github_document
---
```{r, warning=F, message=F}
library(here)
library(tidyverse)
library(tidylog)
library(ggplot2)
library(RColorBrewer)
library(ggdist)
library(plotly)
library(ggpubr)
library(lme4)
library(visreg)
library(DHARMa)

theme_set(theme_minimal())
```

Read data:
```{r}
plant.eff <- readRDS(here("data/Effectiveness_plants.rds"))
plant.eff.sd <- readRDS(here("data/Effectiveness_plants_no_pred.rds"))
```

Also plant crop that will be used for point size
```{r}
plant_crop <- readRDS(here("data/effectiveness.rds")) %>% group_by(id_plant) %>%
  summarise(crop=first(initcrop))
```

Calculate median values of RPE and SDE per plant (to later plot one point per plant)
```{r}
plant.eff.sum <- plant.eff %>%
  group_by(id_plant) %>%
  summarise(median.sde=median(plant_sde), median.rpe=median(plant_rpe)) %>%
  left_join(plant_crop)

plant.eff.sum.sd <- plant.eff.sd %>%
  group_by(id_plant) %>%
  summarise(median.sde=median(plant_sde), median.rpe=median(plant_rpe)) %>%
  left_join(plant_crop)
```


## Calculate RPE~SDE model for each of the 1000 posterior distributions

I create 2 data frames, one that will contain the 1000 slopes and the intercepts, and other that will contain the observed median "x values (i.e. median.rpe value for each plant, n=79) and the predicted "y" value for each of the 1000 model.

I will call "pdp" each posterior probability rendered using bayes
```{r}
effcor <- data.frame()
effcor.predict <- data.frame(x=log(plant.eff.sum$median.rpe))
pdp=c(1:1000)

for (i in 1:1000){ 
  plant.eff.pdp <- plant.eff %>% filter(pdp == pdp[[i]]) 
  model <- lm(log(plant_sde)~log(plant_rpe), data=plant.eff.pdp)
  a <- cbind(pdp = pdp[[i]], t(model$coefficients))
  effcor <- rbind(effcor, a)
  y <- log(plant.eff.sum$median.rpe)*model$coefficients[2]+model$coefficients[1]
  y <- as.data.frame(y)
  names(y)[1] <- paste0("y",pdp[[i]])
  effcor.predict <- cbind(effcor.predict, y)
}
```

Summary of the last model (repetition number = 1000) 
```{r}
summary(model)
simulateResiduals(fittedModel = model, plot = T)
visreg(model)
```

Slope distribution:
```{r}
hist(effcor$`log(plant_rpe)`)
```

Mean slope and mean intercept:
```{r}
mean(effcor$`log(plant_rpe)`)
mean(effcor$`(Intercept)`)
```

The mean slope "0.834" is and mean intercept is "-7.48"

```{r}
write_rds(effcor, here("data/Effectiveness_correlation_slopes.rds"))
write_rds(effcor.predict, here("data/Effectiveness_correlation_slope_CI.rds"))
```

```{r}
# effcor <- readRDS(here("data/Effectiveness_correlation_slopes.rds"))
# effcor.predict <- readRDS(here("data/Effectiveness_correlation_slope_CI.rds"))
```


Plot all 1000 posterior results ("pdps") per plant (i.e. n points= 79000) and slopes:
```{r}
ggplot(plant.eff, aes(x=log(plant_rpe), y=log(plant_sde))) + 
  #ggplot(subset(plant.eff, id_plant=="376"), aes(x=log(plant_rpe), y=log(plant_sde))) + 
  geom_point(fill="#023858", pch=21, color="white") +
  labs(x="Fruit energetic supply (Plant reward)", y="Seedlings recruited (Animal reward)") +
  geom_abline(data=effcor, aes(slope=`log(plant_rpe)`, intercept=`(Intercept)`), 
              alpha=0.1, color="grey30") +
  geom_abline(intercept=mean(effcor$`(Intercept)`), slope=mean(effcor$`log(plant_rpe)`),
              color="#023858", size=1.3) +
  labs(x="Plant reward: total energetic supply (kJ)", y="Animal reward: total seedlings recruited") +
  scale_x_continuous(limits=c(log(0.01), log(100000)),
                     breaks=c(0, log(0.01), log(0.1), log(1), log(10), log(100), log(1000), 
                              log(10000), log(100000)),
                     labels = function(x) format(exp(x))) +
  scale_y_continuous(limits=c(log(0.00001), log(100)),
                     breaks=c(0, log(0.00001), log(0.0001), log(0.001), log(0.01), log(0.1), 
                              log(1), log(10), log(100)),
                     labels = function(x) format(exp(x), scientific=F))

#ggsave(filename=here("manuscript/figures/Fig.SG.1_RPE_SDE_cor_distribution.pdf"), width=10, height=6)
```

Calculate Confidence Interval 100% and 90% for the 1000 slopes
```{r}
CI <- effcor.predict %>% pivot_longer(names_to="pdp", cols=c(y1:y1000)) %>%
  mutate(x=as.factor(x)) %>%
  group_by(x) %>%
  summarise(ymin=min(value), ymax=max(value),
            ymin.90=quantile(value, probs=.05), ymax.90=quantile(value, probs=.95)) %>%
  mutate(x=as.numeric(as.character(x)))

ggplot(CI, aes(x=x)) + geom_ribbon(aes(ymax=ymax, ymin=ymin), color="red", fill="red", alpha=0.4) +
  geom_ribbon(aes(ymax=ymax.90, ymin=ymin.90), color="blue", fill="blue", alpha=0.4)
```

```{r}
ggplot(data=plant.eff.sum, aes(x=log(median.rpe), y=log(median.sde), size=crop)) + 
  labs(x="Fruit energetic supply (Plant reward)", y="Seedlings recruited (Animal reward)") +
  theme(legend.position = "bottom") + 
  geom_ribbon(inherit.aes=F, data=CI, aes(x=x, ymax=ymax.90, ymin=ymin.90), fill="grey50", alpha=0.4) +
  geom_abline(intercept=mean(effcor$`(Intercept)`), slope=mean(effcor$`log(plant_rpe)`),
              color="grey50") +
  geom_point(fill="grey40", pch=21, color="white") +
  scale_size_continuous(range = c(2, 8)) +
  scale_x_continuous(limits=c(log(5), log(10000)),
                     breaks=c(0, log(1), log(10), log(100), log(1000), log(10000)),
                     labels = function(x) format(exp(x))) +
  scale_y_continuous(limits=c(log(0.0001), log(100)),
                     breaks=c(0, log(0.001), log(0.01), log(0.1), log(1), log(10)),
                     labels = function(x) format(exp(x), scientific=F)) +
  labs(caption="Shaded area shows 90% CI for slope")
```


#### SAME PROCCESS BUT REMOVING SEED PREDATORS
```{r}
effcor.sd <- data.frame()
effcor.predict.sd <- data.frame(x=log(plant.eff.sum.sd$median.rpe))
pdp=c(1:1000)

for (i in 1:1000){ 
  plant.eff.pdp <- plant.eff.sd %>% filter(pdp == pdp[[i]]) 
  model.sd <- lm(log(plant_sde)~log(plant_rpe), data=plant.eff.pdp)
  a <- cbind(pdp = pdp[[i]], t(model.sd$coefficients))
  effcor.sd <- rbind(effcor.sd, a)
  y <- log(plant.eff.sum.sd$median.rpe)*model.sd$coefficients[2]+model.sd$coefficients[1]
  y <- as.data.frame(y)
  names(y)[1] <- paste0("y",pdp[[i]])
  effcor.predict.sd <- cbind(effcor.predict.sd, y)
}
```


Summary of the last model (repetition number 1000) and slope distribution
```{r}
summary(model.sd)
visreg(model.sd)
hist(effcor.sd$`log(plant_rpe)`)
```
Now the slope distribution is more normal?

```{r}
mean(effcor.sd$`log(plant_rpe)`)
mean(effcor.sd$`(Intercept)`)
```
The mean slope is very similar "0.850" (even bit higher) but the mean intercept is lower "-7.22"

```{r}
write_rds(effcor.sd, here("data/Effectiveness_correlation_slopes_sd.rds"))
write_rds(effcor.predict.sd, here("data/Effectiveness_correlation_slope_CI_sd.rds"))
```

Calculate Confidence Interval 90% for the 1000 slopes
```{r}
CI.sd <- effcor.predict.sd %>% pivot_longer(names_to="pdp", cols=c(y1:y1000)) %>%
  mutate(x=as.factor(x)) %>%
  group_by(x) %>%
  summarise(ymin=min(value), ymax=max(value),
            ymin.90=quantile(value, probs=.05), ymax.90=quantile(value, probs=.95)) %>%
  mutate(x=as.numeric(as.character(x)))
# 
# ggplot(CI.sd, aes(x=x)) + geom_ribbon(aes(ymax=ymax, ymin=ymin), color="red", fill="red", alpha=0.4) +
#   geom_ribbon(aes(ymax=ymax.90, ymin=ymin.90), color="blue", fill="blue", alpha=0.4)
```

```{r}
ggplot(data=plant.eff.sum.sd, aes(x=log(median.rpe), y=log(median.sde), size=crop)) + 
  labs(x="Fruit energetic supply (Plant reward)", y="Seedlings recruited (Animal reward)") +
  theme(legend.position = "bottom") + 
  geom_ribbon(inherit.aes=F, data=CI.sd, aes(x=x, ymax=ymax.90, ymin=ymin.90), fill="grey50", alpha=0.4) +
  geom_abline(intercept=mean(effcor.sd$`(Intercept)`), slope=mean(effcor.sd$`log(plant_rpe)`),
              color="grey50") +
  geom_point(fill="grey40", pch=21, color="white") +
  scale_size_continuous(range = c(2, 8)) +
  scale_x_continuous(limits=c(log(5), log(10000)),
                     breaks=c(0, log(1), log(10), log(100), log(1000), log(10000)),
                     labels = function(x) format(exp(x))) +
  scale_y_continuous(limits=c(log(0.0001), log(100)),
                     breaks=c(0, log(0.001), log(0.01), log(0.1), log(1), log(10)),
                     labels = function(x) format(exp(x), scientific=F)) +
  labs(caption="Shaded area shows 90% CI for slope
       Note: Seed Predators removed")
```



**Compare graphs with and without seed predators**
```{r}
ggplot(data=plant.eff.sum, aes(x=log(median.rpe), y=log(median.sde), size=crop)) + 
  labs(x="Fruit energetic supply (Plant reward)", y="Seedlings recruited (Animal reward)") +
  theme(legend.position = "none") + 
  geom_abline(intercept=mean(effcor$`(Intercept)`), slope=mean(effcor$`log(plant_rpe)`),
              color="grey20") +
  geom_point(fill="grey40", pch=21, color="white") +
  geom_abline(data=plant.eff.sum.sd, aes(x=log(median.rpe), y=log(median.sde)), 
              intercept=mean(effcor.sd$`(Intercept)`), slope=mean(effcor.sd$`log(plant_rpe)`),
              color="royal blue") + 
  geom_point(data=plant.eff.sum.sd, aes(x=log(median.rpe), y=log(median.sde), size=crop),
             fill="light blue", pch=21, color="white") +
  scale_size_continuous(range = c(2, 8)) +
  scale_x_continuous(limits=c(log(5), log(10000)),
                     breaks=c(0, log(1), log(10), log(100), log(1000), log(10000)),
                     labels = function(x) format(exp(x))) +
  scale_y_continuous(limits=c(log(0.0001), log(100)),
                     breaks=c(0, log(0.001), log(0.01), log(0.1), log(1), log(10)),
                     labels = function(x) format(exp(x), scientific=F)) +
  labs(caption="Shaded area shows 90% CI for slope
       Note: Seed Predators removed in blue")
```

```{r}
a <- plant.eff.sum %>% mutate(type="all")
b <- plant.eff.sum.sd %>% mutate(type="only_sd") 
merged <- rbind(a,b)

ggplot(merged, aes(x=log(median.rpe), y=log(median.sde), group=id_plant, color=type)) + 
  labs(x="Fruit energetic supply (Plant reward)", y="Seedlings recruited (Animal reward)") +
  geom_point() + geom_line() + scale_color_manual(values=c("grey20","light blue")) + 
  theme(legend.position = "bottom")
```

The seedlings recruited do not really change, but the offered fruit energy increases because predators are consuming and not giving anything.


## CORRELATION COEFFICIENTS
```{r}
cor <- data.frame()
pdp=c(1:1000)

for (i in 1:1000){ 
  plant.eff.pdp <- plant.eff %>% filter(pdp == pdp[[i]]) 
  model <- cor.test(log(plant.eff.pdp$plant_rpe), log(plant.eff.pdp$plant_sde))
  a <- cbind(pdp = pdp[[i]], cor=model$estimate, CI=t(model$conf.int))
  cor <- rbind(cor, a)
}

mean(cor$cor)
mean(cor$V3)
mean(cor$V4)
```
