---
title: "Fig_RPE_SDE_correlation"
output: github_document
---

Libraries:
```{r}
library(here)
library(lme4)
library(visreg)
library(tidyverse)
library(visreg)
library(DHARMa)
library(ggplot2)
library(scales)

theme_set(theme_minimal())
```

Read data:

```{r}
plant.eff <- readRDS(here("data/Effectiveness_plants.rds"))
```

Calculate median values of RPE and SDE per plant (to later plot one point per plant)
```{r}
plant.eff.sum <- plant.eff %>%
  group_by(id_plant) %>%
  summarise(median.sde=median(plant_sde), median.rpe=median(plant_rpe))
```

Extract different plant traits to reflect on the point size and color
```{r}
effect <- readRDS(here("data/effectiveness.rds"))

#Area, initial crop size, Median pulp and Median Fruit quality
plant_traits <- effect %>% group_by(id_plant) %>% 
  summarise(pulp=median(pulp_weight), initcrop=first(initcrop),
            fruit.qlc=median(p.reward),
            area=median(area_new), bird_diversity=length(unique(bird_sp)))

#Merge all "point" traits
plant.eff.sum2 <- plant.eff.sum %>% left_join(plant_traits) %>% filter(!is.na(pulp))
```


Calculate Confidence Interval for the 1000 slopes.

1st. Read slopes data (one slopes for each of the 1000 posterior distributions probability - "pdp")
```{r}
effcor <- readRDS(here("data/Effectiveness_correlation_slopes.rds"))
effcor.predict <- readRDS(here("data/Effectiveness_correlation_slope_CI.rds"))
```

2nd. Calculate CI
```{r}
CI <- effcor.predict %>% pivot_longer(names_to="pdp", cols=c(y1:y1000)) %>%
  mutate(x=as.factor(x)) %>%
  group_by(x) %>%
  summarise(ymin=min(value), ymax=max(value),
            ymin.90=quantile(value, probs=.05), ymax.90=quantile(value, probs=.95), 
            ymin.50=quantile(value, probs=.25), ymax.50=quantile(value, probs=.75)) %>%
  mutate(x=as.numeric(as.character(x)))
```

PLOT where point color reflects bird biodiversity
```{r, fig.width=10, fig.height=6}
ggplot(data=plant.eff.sum2, aes(x=log(median.rpe), y=log(median.sde), size=initcrop, fill=bird_diversity)) +
  labs(x="Plant reward: total energetic supply (kJ)", y="Animal reward: total seedlings recruited") +
  geom_ribbon(inherit.aes=F, data=CI, aes(x=x, ymax=ymax.50, ymin=ymin.50), fill="grey50", alpha=0.4) +
  geom_abline(intercept=mean(effcor$`(Intercept)`), slope=mean(effcor$`log(plant_rpe)`),
              color="grey50") +
  geom_point(color="grey")+
  geom_point(pch=21, color="white") +
  scale_fill_gradient(low="#ffeda0", high="#e34a33") +
  scale_size_continuous(name="Crop size", range = c(2, 9), breaks = c(1000, 10000, 50000, 100000), 
                        labels = function(x) format(x, scientific = F)) +
  #theme(legend.position = c(0.90, 0.2)) + 
  #theme(legend.position = "bottom") + 
  scale_x_continuous(limits=c(log(5), log(10000)),
                     breaks=c(0, log(1), log(10), log(100), log(1000), log(10000)),
                     labels = function(x) format(exp(x))) +
  scale_y_continuous(limits=c(log(0.0001), log(10)),
                     breaks=c(0, log(0.0001), log(0.001), log(0.01), log(0.1), log(1), log(10)),
                     labels = function(x) format(exp(x), scientific=F)) +
  labs(caption="Axes in log scale.
       Shaded area shows 50% CI for slope")

#ggsave(filename=here("manuscript/figures/Fig.4_RPE_SDE_cor_plant.pdf"), width=10, height=6)
```

