---
title: "GLM effectiveness"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=F, message=F}
library(here)
library(lme4)
library(glmmTMB)
library(DHARMa)
library(tidyverse)
suppressPackageStartupMessages(library(tidylog))
library(ggplot2)
library(ggpubr)
library(visreg)
```

List non-legitimate dispersers:
```{r}
nld <- c("Chloris chloris", "Cyanistes caeruleus", "Fringilla coelebs","Parus major",
                    "Pyrrhula pyrrhula","Coccothraustes coccothraustes", "Columba palumbus")
```

Load fruit consumption data:
```{r}
crop.consumed.wide <- readRDS(here("data-raw/SDE_QTY_crop_consumed_wide.rds")) %>% 
  mutate(id_plant=as.factor(id_plant)) 
```

Load effectiveness dataset, where we extract plant traits:
```{r}
eff <- readRDS(here("data/effectiveness.rds")) %>%
  mutate(qty.fru.manipulated=as.numeric(qty.fru.manipulated))
```

Summarize plant data
```{r}
plant.traits <- eff %>% group_by(id_plant) %>%
  summarise(site=first(site), pulp.median=median(pulp_weight),
            pulp.sd=sd(pulp_weight), initcrop=first(initcrop), 
            area=first(area_new))
```

Summarize consumption data per individual plant by grouping bird consumption (i.e. 79 observations, not 80 because plant-371 doesn't have data pulp mass). We extract one value for all birds consumption (fru.cons.all) and one value for only legitimate dispersers (fru.cons.disp).
```{r}
rpe.plants <- crop.consumed.wide %>% 
  dplyr::select(-all_of(nld)) %>%
  mutate(fru.cons.per.bird = rowSums(across('Cyanopica cooki':'Turdus viscivorus'), na.rm = T)) %>%
  group_by(id_plant) %>%
  summarise(fru.cons.all=median(fruits.consumed.total),
            fru.cons.disp=median(fru.cons.per.bird)) %>%
  left_join(plant.traits) %>% 
  filter(!is.na(pulp.median)) #remove plant 371 (no fruit data)
```

Round fruits consumed to convert into count data:
```{r}
rpe.plants <- rpe.plants %>% 
  mutate(fru.cons.all.r=round(fru.cons.all)) %>%
  mutate(fru.cons.disp.r=round(fru.cons.disp))
```

Response variable distribution:
```{r}
hist(rpe.plants$fru.cons.disp.r, breaks=50)
```


**MODEL FOR TOTAL CONSUMPTION PER PLANT**

Data is aggregates all bird species consumption by plant
```{r}
m1 <- glmmTMB(fru.cons.disp.r ~ log(initcrop) + log(pulp.median) + log(area) + site, family = "nbinom2", data=rpe.plants)

summary(m1)

m1.output <- simulateResiduals(fittedModel = m1, plot = T)

visreg(m1,  scale="response")
visreg(m1, xvar = "pulp.median", by="site", scale="response")
```
Pulp mean has a positive effect, but with considerable uncertainty.

```{r}
ggplot(rpe.plants, aes(fru.cons.disp.r, pulp.median)) + geom_point() + stat_cor() + theme_bw()
```

There is a clear positive effect of area and initcrop on the total number of fruits consumed.
Initcrop and area are correlated but not colinear, there is big dispersion in their relation.
```{r}
ggplot(rpe.plants, aes(area, initcrop)) + geom_point() + stat_cor() + theme_bw()
```

Relation crop ~ pulp (dot size=consumption)
```{r}
ggplot(data=rpe.plants, aes(x=initcrop, y=pulp.median, size=fru.cons.disp.r, color=site)) + 
  geom_point() + theme_bw() +
  scale_size_continuous(name="Consumption")
```

--------------------------------------------------------------------------------

```{r session, echo= T}
sessioninfo::session_info()
```