---
title: "Null_model_asym_sig"
output: github_document
---

```{r, message=F}
library(here)
library(tidyverse)
library(bipartite)

theme_set(theme_minimal())
```


List of seed predators:

```{r}
sp <- c("Chloris chloris", "Fringilla coelebs", "Pyrrhula pyrrhula", "Coccothraustes coccothraustes", "Columba palumbus")
```

------------------------------------------------------------------------

To estimate the significance of the observed asymmetry values, we
generate null webs based on our fruit consumption findings, and see if
when species interact “randomly”, they would provide similar asymmetry
distributions. We will use 3 different null models: patefield, swap and
vazquez.

To do the web reshuffling, we need to work with the consumption network
(i.e., the QTC). Then, given the constant quality values for the plants
and dispersers, we calculate di novo the value of RPE and SDE for each
of the generated null webs

1st. Read the raw interaction matrix based on the number of fruits
consumed (QTC). We remove all interactions with Coccothraustes, since
the number of fruits consumed is 0 when rounded, so there is no
dependence (this 0 value creates conflict with the Vazquez null model,
so it is removed from all analyses at the beginning).

``` {r}
int.mat <- read_csv(here("data/consumed_fruit_mat.csv")) %>% column_to_rownames(var="id_plant") %>% 
  round() %>% select(-Coc.coc)

int.mat[is.na(int.mat)] <- 0
```

2nd. Read data for extracting quality (QLC)

``` {r}
eff <- readRDS(here("data/effectiveness.rds"))
```


**Bird quality:** Extract bird quality, create birds acronyms and remove
coccotharustes

``` {r}
bird_quality <- eff %>% group_by(bird_sp) %>% 
  summarise(sde.qlc=median(sde.qlc)) %>% 
  separate(bird_sp, into = c("genus", "sp"), remove = FALSE) %>% 
  mutate(bird_sp_short = paste0(str_sub(genus, end = 3), ".", str_sub(sp, end = 3))) %>% 
  mutate(bird_sp_short = ifelse(bird_sp_short == "NA.NA", NA, bird_sp_short)) %>% 
  select(-genus, -sp) %>% 
  relocate(bird_sp_short, .after = bird_sp) %>%
  filter(bird_sp_short!="Coc.coc")
```


**Plant quality:** Extract quality for pulp and seed, as different for
seed predators and pulp consumers.

``` {r}
plant_quality <- eff %>% group_by(id_plant) %>% 
  summarise(pulp.qlc=median(p.reward), seed.qlc=median(s.reward))
```


Calculate RPE and SDE with the qtc matrix and qlc values:

``` {r}
long.int <- int.mat %>% rownames_to_column("id_plant") %>%
  pivot_longer(cols = c(2:27), names_to="bird_sp_short", values_to="qtc") %>%
  left_join(bird_quality) %>% left_join(plant_quality) %>%
  mutate(rpe.qlc=ifelse(bird_sp %in% sp, seed.qlc, pulp.qlc)) %>%
  mutate(sde=qtc*sde.qlc, rpe=qtc*rpe.qlc)
```

Calculate **observed** dependencies and asymmetry:

``` {r}
eff.dep <- long.int %>% 
  mutate(sde=ifelse(qtc==0, NA, sde),
         rpe=ifelse(qtc==0, NA, rpe)) %>%
  group_by(id_plant) %>%
  mutate(sde.plant=sum(sde, na.rm=T)) %>%
  mutate(sde.dep=sde/sde.plant) %>%
  ungroup() %>%
  group_by(bird_sp) %>%
  mutate(rpe.bird=sum(rpe, na.rm = T)) %>%
  mutate(rpe.dep=rpe/rpe.bird) %>%
  ungroup() %>%
  mutate(asy.num=sde.dep-rpe.dep) %>%
  rowwise() %>%
  mutate(asy.denom=ifelse(is.na(sde.dep), NA, max(sde.dep, rpe.dep))) %>%
  mutate(asymmetry=asy.num/asy.denom) %>%
  mutate(int=paste0(id_plant, "-", bird_sp_short))
```

**Null models**
(<https://fukamilab.github.io/BIO202/09-B-networks.html>)

A common binary null model:

-   ‘shuffle.web’ - shuffles cells around the interaction matrix, keeps
    nothing constant

Common weighted null models:

-   ‘r2dtable’ - changes cell values while keeping row and column sums
    constant

-   ‘vaznull’ - changes cell values while keeping connectance (number of
    links in the matrix) constant
    

# PATEFIELD NULL MODELS

Generate Patefield’s null models for QTC web

``` {r echo=T, eval=F}
nulls <- nullmodel(round(int.mat), N=1000, method="r2dtable")

#write_rds(nulls, here("data/null_models_patefield.rds"))
```

Calculate dependence and asymmetry values for Patefield null webs

``` {r echo=T, eval=F}
null.as<-eff.dep$int

for (i in 1:1000){ 
  subset.nulls <- nulls[[i]]
  colnames(subset.nulls) <- colnames(int.mat)
  rownames(subset.nulls) <- rownames(int.mat)
  long.int.null <- as.data.frame(subset.nulls) %>% 
    rownames_to_column("id_plant") %>%
    pivot_longer(cols = c(2:27), names_to="bird_sp_short", values_to="qtc") %>%
    left_join(bird_quality) %>% left_join(plant_quality) %>%
    mutate(rpe.qlc=ifelse(bird_sp %in% sp, seed.qlc, pulp.qlc)) %>%
    mutate(sde=qtc*sde.qlc, rpe=qtc*rpe.qlc)
  eff.dep.null <- long.int.null %>% 
    mutate(sde=ifelse(qtc==0, NA, sde),
         rpe=ifelse(qtc==0, NA, rpe)) %>%
    group_by(id_plant) %>%
    mutate(sde.plant=sum(sde, na.rm = T)) %>%
    mutate(sde.dep=sde/sde.plant) %>%
    ungroup() %>%
    group_by(bird_sp) %>%
    mutate(rpe.bird=sum(rpe, na.rm = T)) %>%
    mutate(rpe.dep=rpe/rpe.bird) %>%
    ungroup() %>%
    mutate(asy.num=sde.dep-rpe.dep) %>%
    rowwise() %>%
    mutate(asy.denom=ifelse(is.na(sde.dep), NA, max(sde.dep, rpe.dep))) %>%
    mutate(asymmetry=asy.num/asy.denom)
  names(eff.dep.null)[17] <- paste0("as", i)
  null.as <- cbind(null.as, eff.dep.null[17])
}

write_rds(null.as, here("data/asymm_null_models_patefield.rds"))
```

If not run loop, then load data:

``` {r}
null.as <- readRDS(here("data/asymm_null_models_patefield.rds"))
```

Plot observed asymmetry distribution vs. the distribution of the 1000
null patefield generated asymmetries:

``` {r}
patefield <- data.frame(as.list(null.as)) %>% 
  pivot_longer(cols = c(-1), values_to = "asymmetry", names_to="model")

ggplot(eff.dep, aes(x=asymmetry, y=..count..)) + 
  geom_line(inherit.aes=F, data=patefield, aes(x=asymmetry, group=model, y=..count..), 
            stat="density", bw=0.04,  color="#a6bddb", alpha=.05) + 
  geom_line(stat="density", bw=0.04) + 
  scale_x_continuous(limits=c(-1.1,1.1))
```

# VAZQUEZ NULL MODELS

Generate Vazquez null models for QTC web

``` {r echo=T, eval=F}
nulls.vaz <- nullmodel(ceiling(int.mat), N=1000, method="vaznull")

#write_rds(nulls.vaz, here("data/null_models_vazquez.rds"))
```

Calculate dependence and asymmetry values for vazquez null webs

``` {r echo=T, eval=F}
null.as.vaz <- eff.dep$int

for (i in 1:1000){ 
  subset.nulls <- nulls.vaz[[i]]
  colnames(subset.nulls) <- colnames(int.mat[-27])
  rownames(subset.nulls) <- rownames(int.mat)
  long.int.null <- as.data.frame(subset.nulls) %>% 
    rownames_to_column("id_plant") %>%
    pivot_longer(cols = c(2:27), names_to="bird_sp_short", values_to="qtc") %>%
    left_join(bird_quality) %>% left_join(plant_quality) %>%
    mutate(rpe.qlc=ifelse(bird_sp %in% sp, seed.qlc, pulp.qlc)) %>%
    mutate(sde=qtc*sde.qlc, rpe=qtc*rpe.qlc)
  eff.dep.null <- long.int.null %>% 
    mutate(sde=ifelse(qtc==0, NA, sde),
         rpe=ifelse(qtc==0, NA, rpe)) %>%
    group_by(id_plant) %>%
    mutate(sde.plant=sum(sde, na.rm = T)) %>%
    mutate(sde.dep=sde/sde.plant) %>%
    ungroup() %>%
    group_by(bird_sp) %>%
    mutate(rpe.bird=sum(rpe, na.rm = T)) %>%
    mutate(rpe.dep=rpe/rpe.bird) %>%
    ungroup() %>%
    mutate(asy.num=sde.dep-rpe.dep) %>%
    rowwise() %>%
    mutate(asy.denom=ifelse(is.na(sde.dep), NA, max(sde.dep, rpe.dep))) %>%
    mutate(asymmetry=asy.num/asy.denom)
  names(eff.dep.null)[17] <- paste0("as", i)
  null.as.vaz <- cbind(null.as.vaz, eff.dep.null[17])
}

write_rds(null.as.vaz, here("data/asymm_null_models_vazquez.rds"))
```

If not run loop, then load data:

``` {r}
null.as.vaz <- readRDS(here("data/asymm_null_models_vazquez.rds"))
```

Plot observed asymmetry distribution vs. the distribution of the 3 null
model generated asymmetries (Patefield in blue, swap in red and vazquez
in orange):

``` {r}
vaz <- data.frame(as.list(null.as.vaz)) %>% 
  pivot_longer(cols = c(-1), values_to = "asymmetry", names_to="model")

ggplot(eff.dep, aes(x=asymmetry, y=..count..)) + 
  geom_line(inherit.aes=F, data=patefield, aes(x=asymmetry, group=model, y=..count..), 
                stat="density", alpha=0.05, color="#a6bddb", bw=0.04) +
  geom_line(inherit.aes=F, data=vaz, aes(x=asymmetry, group=model, y=..count..), 
                stat="density", alpha=0.05, color="#fc8d62", bw=0.04) +
  geom_line(stat="density", bw=0.04) + scale_x_continuous(limits=c(-1.1,1.1)) + theme_bw()

#ggsave(here("manuscript/figures/Fig.SH.2_Null_model_significance.pdf"), width = 6, height = 5)
```


