---
title: "Null model dimensions"
output: github_document
---

``` {r message=F}
library(here)
library(tidyverse)
library(tidylog)
library(ggplot2)
library(RColorBrewer)
library(ggdist)
library(plotly)
library(ggpubr)

theme_set(theme_minimal())
```

Read SDE and RPE values:

``` {r}
effect <- readRDS(here("data/effectiveness.rds"))
```

Summarise data to calculate median RPE and SDE:

``` {r}
effect_sum <- effect %>%
  mutate(sde.eff=qty.fru.manipulated*sde.qlc, rpe.eff=qty.fru.consumed*fruit.reward) %>%
  group_by(id_plant, bird_sp) %>%
  summarise(median.sde=median(sde.eff), median.rpe=median(rpe.eff),
            initcrop=first(initcrop), area=first(area_new))
```

In order to do simulations we need to work with the RPE and SDE values,
for which I generate one matrix with SDE values and another matrix with
RPE values, and merge together into one (where rows are plants and each
animals has two columns, one with RPE and on with SDE). In this way we
can later do subsamplings on the number of plants (i.e. number of rows).

``` {r}
sde.mat <- effect_sum %>%
    mutate(bird_sp=paste0(bird_sp,"_SDE")) %>%
    select(id_plant, bird_sp, median.sde) %>%
    pivot_wider(names_from=bird_sp, values_from=median.sde) %>% 
    ungroup()
```

``` {r}
rpe.mat <- effect_sum %>% 
    mutate(bird_sp=paste0(bird_sp,"_RPE")) %>%
    select(id_plant, bird_sp, median.rpe) %>%
    pivot_wider(names_from=bird_sp, values_from=median.rpe) %>%
    ungroup()
```

``` {r}
mat.full <- full_join(sde.mat, rpe.mat) %>% filter(id_plant!=371)
```

Now I subsample 20, 40 and 60 rows (n=1000 repetitions for each subsampling cut) - this loop takes some time (~15min on my computer).
Can just read it at the end of the chunk.

``` {r echot=T, eval=F}
asy.permutations <- data.frame()

for (k in c(20, 40, 60)){
  for (i in 1:1000){
sampling <- mat.full %>%
  slice_sample(n=k, replace=F) %>%
  pivot_longer(cols=c("Chloris chloris_SDE":"Coccothraustes coccothraustes_RPE"),
               names_to="bird_sp", values_to="effectiveness") %>%
  mutate(type=str_sub(bird_sp, start= -3)) %>%
  mutate(bird_sp=str_sub(bird_sp, end= -5)) %>%
  pivot_wider(names_from=type, values_from=effectiveness) %>%
  group_by(id_plant) %>%
  mutate(sde.plant=sum(SDE, na.rm=T)) %>%
  mutate(sde.dep=SDE/sde.plant) %>%
  ungroup() %>%
  group_by(bird_sp) %>%
  mutate(rpe.bird=sum(RPE, na.rm = T)) %>%
  mutate(rpe.dep=RPE/rpe.bird) %>%
  ungroup() %>%
  rowwise() %>%
  mutate(asy.num=sde.dep-rpe.dep) %>%
  mutate(asy.denom=ifelse(!is.na(asy.num), max(sde.dep, rpe.dep, na.rm = T), NA)) %>%
  mutate(asymmetry=asy.num/asy.denom) %>%
  filter(!is.na(asymmetry))
ele <- sampling  %>% select(asymmetry) %>% mutate(permutation=i, size=k)
asy.permutations <- rbind(asy.permutations, ele)
  }
}

write_rds(asy.permutations, here("data/null_model_matrix_dimensions.rds"))
```

Read output of loop:
``` {r}
asy.permutations <- readRDS(here("data/null_model_matrix_dimensions.rds"))
```

Specify matrix size as treatment (convert from numeric to character).

``` {r}
asy.permutations <- as.data.frame(asy.permutations) %>% 
  mutate(size2 = ifelse(size==20, "20 plants", "other"),
         size2 = ifelse(size==40, "40 plants", size2),
         size2 = ifelse(size==60, "60 plants", size2))
```

Load observed dependence and asymmetry values data:

``` {r}
dep <- readRDS(here("data/Dependency_distribution.rds"))
```

Summarise observed data:

``` {r}
dep.sum <- dep %>% group_by(id_plant, bird_sp) %>%
  summarise(rpe.animal=median(rpe.animal.1.pdp), sde.plant=median(sde.plant.1.pdp),
            rpe=median(rpe.eff), sde=median(sde.eff),
            animal.dep=median(rpe.dep), plant.dep=median(sde.dep),
            asymmetry=median(asymmetry)) %>% mutate(size=80)
```

Plot all asymmetries distribution for different matrix sizes.

``` {r}
ggplot(asy.permutations, aes(x=asymmetry, group=permutation)) + 
  geom_line(stat="density", bw=0.04, alpha=.05, colour="#a6bddb") + 
  geom_density(inherit.aes=F, data=dep.sum, aes(x=asymmetry), bw=0.04) +
  facet_grid(~size2) + theme(legend.position = "none") + scale_x_continuous(limits=c(-1.1,1.1)) +
  theme_bw()
  # labs(caption = "Black line represents observed asymmetry. 
  #      There are 1000 repetitions for matrices of different dimensions (20, 40 and 60 rows)")

#ggsave(here("manuscript/figures/Fig.SH.1_Null_model_dims.pdf"), width = 7, height = 5)
```

Observed asymmetry distribution: Shaded lines represent asymmetry for
each of the 1000 posterior distribution probabilities, and the dark line the median asymmetry.

``` {r}
obs.dep.selec <- dep %>% select(asymmetry, pdp) %>% mutate(size=80)

ggplot(obs.dep.selec, aes(x=asymmetry, group=pdp)) + 
  geom_line(stat="density", bw=0.04, alpha=.05, color="grey70") + 
  geom_line(inherit.aes=F, data=dep.sum, aes(x=asymmetry), stat="density", bw=0.04)
  # labs(caption = "Black line represents asymmetry calculated using median values. 
  #      Thin transparent grey lines represent asymmetry for each of the 1000 posteriors")
```