---
title: "Number of fruits consumed | visit"
date: "`r Sys.Date()`"
output: github_document
---


```{r setup, include=FALSE}
set.seed(888)
library(knitr)
knitr::opts_chunk$set(echo = TRUE, out.width = '100%', 
                      cache = TRUE, message = FALSE)
```

```{r Number-fruits-visit-1 }
library(here)
library(dplyr)
library(ggplot2)
library(brms)
library(DHARMa)

source(here("code", "functions.R"))

theme_set(theme_minimal())
```


Load data

```{r Number-fruits-visit-2 }
nfrudata <- readr::read_csv(here("data", "lentisc_feeding_rates.csv"))
```

As some spp were missing from local dataset, use body mass data from EltonTraits (https://doi.org/10.6084/m9.figshare.c.3306933.v1)

```{r Number-fruits-visit-3}
elton <- readr::read_delim(here("data", "BirdFuncDat.txt"), delim = "\t") %>% 
  select(bird_sp = Scientific, bmass.elton = `BodyMass-Value`) %>% 
  mutate(bird_sp = ifelse(bird_sp == "Parus caeruleus", "Cyanistes caeruleus", bird_sp),
         bird_sp = ifelse(bird_sp == "Cyanopica cyanus", "Cyanopica cooki", bird_sp),
         bird_sp = ifelse(bird_sp == "Saxicola torquatus", "Saxicola rubicola", bird_sp),
         bird_sp = ifelse(bird_sp == "Carduelis chloris", "Chloris chloris", bird_sp))

nfrudata <- nfrudata %>% 
  left_join(elton) %>% 
  mutate(log.bodym.Dg = log(bmass.elton/10) - log(2)) # convert to Dg, take log, and center on 2Dg bird

```


## EDA

Number of fruits consumed per species:

```{r Number-fruits-visit-4}

nfrudata %>% 
  group_by(diet, bird_sp) %>% 
  summarise(mean(no_fruits), 
            sd(no_fruits), 
            n()) %>% 
  kable(digits = 2)

nfrudata %>% 
  ggplot() +
  facet_wrap(~bird_sp, scales = "free") +
  geom_histogram(aes(no_fruits))
```


Relationship with body mass:

```{r Number-fruits-visit-5}
sort(round(unique((nfrudata$bodymass))))

nfrudata %>% 
  filter(bird_sp != "Chloris chloris") %>%
  ggplot() +
  geom_point(aes(log(bmass.elton), log(no_fruits)))

```


We're modelling *Chloris chloris* separately as eager seed predator. The remaining species consume typically 2-3-4 fruits/visit.




## Model all species except Chloris

```{r Number-fruits-visit-6 }
nfrudata.redux <- nfrudata %>% 
  filter(bird_sp != "Chloris chloris") 
```


Prior predictive check:

```{r Number-fruits-visit-7 }
# Using negbinomial as there is overdispersion 
# Also, truncated at 1 (no 0 counts in dataset)
nfru.formu <- brmsformula(no_fruits | trunc(lb = 1) ~ (1 | bird_sp) + log.bodym.Dg,
                         family = negbinomial())

get_prior(nfru.formu, data = nfrudata.redux)

# priors (log scale)
# log(2) = 0.7  # 2 fruits/visit as grand mean
priors <- c(
  set_prior("normal(0.7, 0.3)", class = "Intercept"), 
  set_prior("normal(0.5, 0.5)", class = "b"),
  set_prior("normal(0, 0.5)", class = "sd")
)

```

Fit model:

```{r Number-fruits-visit-8 }

nfru.m <- brm(nfru.formu,
         data = nfrudata.redux,
         prior = priors,
         cores = 4,
         iter = 4000,
         refresh = 0, 
         control = list(adapt_delta = 0.95))


pp_check(nfru.m, nsamples = 100)
pp_check(nfru.m, nsamples = 100) + xlim(0,10)

bayestestR::check_prior(nfru.m)

summary(nfru.m, priors = TRUE)

plot(nfru.m)

stancode(nfru.m)

coef(nfru.m)

#DHARMa
nfru.check <- check_brms(nfru.m, integer = TRUE)
plot(nfru.check, form = nfrudata.redux$bird_sp)
testDispersion(nfru.check)

saveRDS(nfru.m, here("analysis", "output", "nfru_visit_model.rds"))

```


### Predict

Add species without feeding observations:

```{r Number-fruits-visit-9}
pfeed.post <- readRDS(here("data", "Pfeed.post.rds"))
pfeed.sp <- sort(unique(pfeed.post$bird_sp))

nfru.sp <- sort(unique(nfrudata$bird_sp))

sp.new <- pfeed.sp[!pfeed.sp %in% nfru.sp]
sp.new

newdat <- data.frame(bird_sp = sort(c(unique(nfrudata.redux$bird_sp), 
                                      sp.new)))

# Add bodymass data for these spp
newdat <- newdat %>% 
  left_join(elton) %>% 
  mutate(log.bodym.Dg = log(bmass.elton/10) - log(2)) # convert to Dg, take log, and center on 2Dg bird

```


```{r Number-fruits-visit-10 }
## Full posterior distributions
nfru.post <- posterior_predict(nfru.m, 
                                newdata = newdat,
                                nsamples = 1000,
                                allow_new_levels = TRUE, 
                               ntrys = 20) %>% 
  as.data.frame() %>%
  setNames(newdat$bird_sp)

nfru.post.long <- nfru.post %>% 
  tidyr::pivot_longer(cols = everything(), names_to = "bird_sp", values_to = "Nfruits") 

nfru.post.long %>% 
  group_by(bird_sp) %>% 
  summarise(mean(Nfruits), sd(Nfruits)) %>% 
  arrange(desc(`mean(Nfruits)`)) %>% 
  kable(digits = 2)

# saveRDS(nfru.post.long, here("analysis", "output", "nfruits.post.long.rds"))

```





-----




## Model *Chloris chloris* independently

```{r Number-fruits-visit-11 }
chloris <- nfrudata %>% 
  filter(bird_sp == "Chloris chloris")

chloris$row <- 1:nrow(chloris)
```


Prior predictive check:

```{r Number-fruits-visit-12 }
# Using negbinomial as there is overdispersion 
# Also, truncated at 1 (no 0 counts in dataset)
nfru.c.formu <- brmsformula(no_fruits | trunc(lb = 1) ~ 1,
                         family = negbinomial())

get_prior(nfru.c.formu, data = chloris)

# priors (log scale)
# exp(2) = 7.4
priors <- c(
  set_prior("normal(2, 0.5)", class = "Intercept")
)


```

Fit model:

```{r Number-fruits-visit-13 }

nfru.c.m <- brm(nfru.c.formu,
         data = chloris,
         prior = priors,
         cores = 4,
         iter = 4000,
         refresh = 0)


pp_check(nfru.c.m, nsamples = 100)
pp_check(nfru.c.m, nsamples = 100) + xlim(0, 50)
bayesplot::ppc_rootogram(y = chloris$no_fruits, 
                         yrep = posterior_predict(nfru.c.m, nsamples = 1000)) + xlim(0,50)

bayestestR::check_prior(nfru.c.m)

summary(nfru.c.m, priors = TRUE)

plot(nfru.c.m)

stancode(nfru.c.m)

# coef(nfru.c.m)
fixef(nfru.c.m)
# ranef(nfru.m)

# mcmc_plot(nfru.m, pars = "id_plant")
# launch_shinystan(nfru.m)

#DHARMa
nfru.c.check <- check_brms(nfru.c.m, integer = TRUE)
testDispersion(nfru.c.check)

saveRDS(nfru.c.m, here("analysis", "output", "nfru_visit_Chloris_model.rds"))

```


### Predict


```{r Number-fruits-visit-14 }
## Full posterior distributions
nfru.c.post <- posterior_predict(nfru.c.m, 
                               newdata = data.frame(bird_sp = "Chloris chloris"),
                               nsamples = 1000,
                               ntrys = 20
                               ) %>% 
  as.data.frame() %>%
  setNames("Nfruits") %>% 
  bind_cols(bird_sp = "Chloris chloris") %>% 
  relocate(bird_sp)

summary(nfru.c.post$Nfruits)

```



## Combine Chloris with other spp

```{r Number-fruits-visit-15}
nfru.post.long <- bind_rows(nfru.post.long, nfru.c.post)

nfru.post.long %>% 
  group_by(bird_sp) %>% 
  summarise(mean(Nfruits), sd(Nfruits), 
            max(Nfruits), q95 = round(quantile(Nfruits, probs = 0.95))) %>% 
  arrange(desc(`mean(Nfruits)`)) %>% 
  kable(digits = 2)

saveRDS(nfru.post.long, here("analysis", "output", "Nfruits_visit_post_long.rds"))
```

