---
title: "SDE QLTY: Probability of microhabitat use"
date: "`r Sys.Date()`"
output: github_document
---


```{r setup, include=FALSE}
set.seed(888)
library(knitr)
knitr::opts_chunk$set(echo = TRUE, out.width = '100%', 
                      cache = TRUE, message = FALSE)
```

```{r }
library(here)
library(readr)
library(dplyr)
library(ggplot2)
suppressPackageStartupMessages(library(tidylog))

source(here("code", "functions.R"))

theme_set(theme_minimal())
```


Load data

```{r}
mhdata <- read_csv(here("data", "mh_combined_barcoding.csv"), guess_max = 10000) %>% 
    rename(sampling_point = id_plant,
         sample_area_m2 = indiv_samp_eff) %>% 
  filter(sample_area_m2 != "out_transect") %>% 
  mutate(sample_area_m2 = as.numeric(sample_area_m2))

```


Keep only droppings w/ Pistacia seeds

```{r}
mhdata.pis <- filter(mhdata, seed_sp1 == "pistacia") 
glimpse(mhdata.pis)
```

```{r}
summarytools::dfSummary(mhdata.pis, plain.ascii = FALSE, style = "grid", valid.col = FALSE)
```


There are 5 MH (PL, FR, NF, PP, OA). PL has 40 sampling points, FR, NF and PP have 15 sampling points each, and OA there are 17 transects of different lengths (made on different days).

Add two OA transects to `mhdata.pis` which didn't have any PL seed:

```{r}
oa.mis <- unique(mhdata$sampling_point)[! unique(mhdata$sampling_point) %in% unique(mhdata.pis$sampling_point)]

oa.areas <- unique(mhdata$sample_area_m2[mhdata$sampling_point %in% oa.mis])

mhdata.pis <- mhdata.pis %>% 
  add_row(sampling_point = oa.mis, mh = "OA", sample_area_m2 = oa.areas, 
          seed_sp1 = "pistacia", seed_n1 = 0)

```


Comparing full dataset w/ Pistacia only:

```{r}

table(mhdata$bird_sp, mhdata$mh)
table(mhdata.pis$bird_sp, mhdata.pis$mh)

round(table(mhdata$bird_sp, mhdata$mh)/nrow(mhdata), digits = 2)
round(table(mhdata.pis$bird_sp, mhdata.pis$mh)/nrow(mhdata.pis), digits = 2)

sum(is.na(mhdata$bird_sp))
sum(is.na(mhdata.pis$bird_sp))

```

Remove bird species that were not observed consuming fruits (*Alectoris rufa*):

```{r}
mhdata.pis <- mhdata.pis %>% 
  filter(bird_sp != "Alectoris rufa" | is.na(bird_sp))
```


Total number of PL seeds collected (not all assigned to specific bird):

```{r}
sum(mhdata.pis$seed_n1)
```


Seed rain per microhabitat (note sampling area has not been homogeneous across MH):

```{r}
mhdata.pis %>% 
  group_by(mh) %>% 
  summarise(nseed = sum(seed_n1)) %>% 
  kable()
```




Seed rain per sampling point and bird:

```{r}
seeds.mh <- mhdata.pis %>% 
  group_by(mh, sampling_point, sample_area_m2) %>% 
  summarise(nseed = sum(seed_n1)) %>% 
  ungroup() %>% 
  mutate(nseed.m2 = nseed / sample_area_m2)

summary(seeds.mh)
table(seeds.mh$mh)

seeds.mh.bird <- mhdata.pis %>% 
  group_by(mh, sampling_point, sample_area_m2, bird_sp) %>% 
  summarise(nseed = sum(seed_n1))

summary(seeds.mh.bird)
```



## MODEL NSEED PER MH


First we model the total seed rain per MH. In a second model below we assign bird species.
This is mainly because bird species has not been identified for all seeds (otherwise we could feed count model to seeds dispersed per MH and bird species).


### Prior Predictive Check


```{r }

library(brms)

Nseed.formu <- brmsformula(nseed | rate(sample_area_m2) ~ mh, 
                            family = negbinomial())

get_prior(Nseed.formu, data = seeds.mh)

## priors (log scale)
# exp(4) = 54 seeds/m2
# exp(3) = 20 seeds/m2
# exp(5) = 148 seeds/m2
# plot(density(rlnorm(1000, 3, 1)))
priors <- c(
  set_prior("normal(3, 2)", class = "Intercept"), 
  set_prior("normal(0, 2)", class = "b")
)
```


Prior predictive check:

```{r eval=FALSE}
Nseed.m <- brm(Nseed.formu,
         data = seeds.mh,
         prior = priors,
         sample_prior = "only",
         cores = 4,
         iter = 2000,
         refresh = 0)

# stancode(Nseed.m)
pp_check(Nseed.m, ndraws = 100) 
summary(Nseed.m, priors = TRUE)
plot(Nseed.m)
```



### Fit model


```{r }
Nseed.m <- brm(Nseed.formu,
         data = seeds.mh,
         prior = priors,
         # sample_prior = "only",
         cores = 4,
         iter = 4000,
         refresh = 0)
```


Check: 

```{r }
library(DHARMa)

pp_check(Nseed.m, ndraws = 100)
bayesplot::ppc_rootogram(y = seeds.mh$nseed, 
                         yrep = posterior_predict(Nseed.m, ndraws = 1000))
# pp_check(Pvisit.m, ndraws = 100, type = "error_scatter_avg")
bayestestR::check_prior(Nseed.m)

summary(Nseed.m, priors = TRUE)

plot(Nseed.m, ask = FALSE)

stancode(Nseed.m)

# coef(Nseed.m)
fixef(Nseed.m)

modelbased::estimate_means(Nseed.m)

#mcmc_plot(Nseed.m, pars = "sampling_point")
# launch_shinystan(Nseed.m)

#DHARMa
Nseed.check <- check_brms(Nseed.m, integer = TRUE)
plot(Nseed.check, form = seeds.mh$mh)
plot(Nseed.check, form = seeds.mh$sampling_point)
testDispersion(Nseed.check)

saveRDS(Nseed.m, here("analysis", "output", "nseed.mh.model.rds"))

```



### Predict

```{r }

## Full posterior distributions
# Mean number of seeds per m2 per MH
Nseed.post <- posterior_predict(Nseed.m, 
                                newdata = data.frame(
                                  mh = unique(seeds.mh$mh),
                                  # sampling_point = seeds.mh$sampling_point,
                                  sample_area_m2 = 1   # seeds per m2
                                  ),
                                re_formula = NA,  # do not include site random effects, only MH
                                ndraws = 1000) %>% 
  as.data.frame() %>%
  setNames(unique(seeds.mh$mh)) 


Nseed.post.long <- Nseed.post %>% 
  tidyr::pivot_longer(cols = everything(), names_to = "mh", values_to = "seeds.m2") %>%
  arrange(mh)

saveRDS(Nseed.post.long, here("analysis", "output", "Nseed.mh.post.rds"))

```


Sum total number of seeds for each iteration to calculate probability of a single seed being dispersed to each MH:

```{r }
Pseed.post <- Nvisit2Pvisit(Nseed.post, tidy = FALSE) %>% 
  as.data.frame() %>% 
  tidyr::pivot_longer(cols = everything(), names_to = "mh", values_to = "Pseed.mh") %>% 
  arrange(mh)

Pseed.post %>% group_by(mh) %>% summarise(mean(Pseed.mh))

saveRDS(Pseed.post, here("analysis", "output", "Pseed.mh.post.rds"))

```





------------------------------------------------------------------

### 2. Pbird: Probability that each bird species deposits seeds in different microhabitats

#### 2.1 Prepare data

Get number of seeds dispersed by each bird species to each microhabitat sampling point. There are quite a few NA (`r sum(seeds.mh.bird$nseed[is.na(seeds.mh.bird$bird_sp)])` out of `r sum(seeds.mh.bird$nseed)`). We remove them as it could be any species, including already identified species, so if we would count the unidentified seeds for the sampling_point total we could be biasing down the frequency of the other species.



```{r }
seeds.mh.sp <- seeds.mh.bird %>% 
  filter(!is.na(bird_sp))
```


Add other unobserved combinations of deposition by bird species on sampling points (assigning nseed = 0):

```{r }
## load bird species observed consuming
bird.consum <- readRDS(here("analysis", "output", "SDE_QTY_Nfruits.rds")) %>% 
  distinct(bird_sp) %>% 
  pull()

bird.consum

all.int <- expand.grid(sampling_point = sort(unique(seeds.mh$sampling_point)), 
                       bird_sp = bird.consum, 
                       stringsAsFactors = FALSE)

unobs <- anti_join(all.int, dplyr::select(seeds.mh.sp, sampling_point, bird_sp)) %>% 
    mutate(nseed = 0) %>% 
  left_join(distinct(dplyr::select(seeds.mh.sp, sampling_point, sample_area_m2)), by = "sampling_point") %>% 
  dplyr::select(mh, sampling_point, sample_area_m2, bird_sp, nseed) %>% 
  mutate(mh = ifelse(stringr::str_starts(sampling_point, "OA"), "OA", mh))

seeds.mh.sp <- full_join(seeds.mh.sp, unobs)
```




Calculate total number of **bird-identified** seeds per MH sampling point: 

```{r }

seeds.mh.id <- seeds.mh.sp %>% 
  group_by(sampling_point) %>% 
  summarise(nseed.total = sum(nseed))

seeds.mh.sp <- seeds.mh.sp %>% 
  left_join(seeds.mh.id, by = "sampling_point") %>% 
  mutate(nseed.bird.prop = nseed/nseed.total) %>% 
  ungroup() %>% 
  arrange(sampling_point, bird_sp)

seeds.mh.sp$row <- 1:nrow(seeds.mh.sp)

```



#### 2.2 Model



```{r }

options(brms.backend = "cmdstanr")

Pbird.formu <- bf(nseed | trials(nseed.total) ~
                    1 + mh + (1 + mh | bird_sp),
                  family = binomial())

get_prior(Pbird.formu, data = seeds.mh.sp)

# priors (logit scale)
# plogis(-6) = 0.0025
# plogis(-7) = 0.0009
# plogis(-8) = 0.00033

# plogis(-2) = 0.12   
# plogis(-3) = 0.047
# plogis(-1) = 0.27
# plogis(-4) = 0.018
# plogis(0) = 0.5

priors <- c(
  prior(normal(-3.3, 1), class = "Intercept"),     #plogis(-3.3) ~~ 1/27 bird spp
  prior(normal(0, 2), class = "sd"),
  prior(lkj(2), class = "cor")  # https://bookdown.org/content/4857/adventures-in-covariance.html
)

```



Prior predictive check:

```{r eval=FALSE}

Pbird.m <- brm(Pbird.formu,
         data = seeds.mh.sp,
         prior = priors,
         sample_prior = "only",
         cores = 4,
         iter = 2000,
         refresh = 0)

# stancode(Nseed.m)
pp_check(Pbird.m, ndraws = 100)
pp_check(Pbird.m, ndraws = 100) + coord_cartesian(xlim = c(0, 5))
summary(Pbird.m, priors = TRUE)
plot(Pbird.m)

```


Fit model:

```{r cache = TRUE}
Pbird.m <- brm(Pbird.formu,
         data = seeds.mh.sp,
         prior = priors,
         cores = 4,
         iter = 4000,
         refresh = 0)
```


Check:

```{r }

pp_check(Pbird.m, ndraws = 100)
pp_check(Pbird.m, ndraws = 100) + coord_cartesian(xlim = c(0, 5), ylim = c(0, 1))
pp_check(Pbird.m, ndraws = 100, type = "bars") + coord_cartesian(xlim = c(0, 5))
# pp_check(Pvisit.m, ndraws = 100, type = "error_scatter_avg")
# suppressWarnings(bayestestR::check_prior(Pbird.m))

summary(Pbird.m, priors = TRUE)

plot(Pbird.m, ask = FALSE)

stancode(Pbird.m)

lapply(coef(Pbird.m), head)
fixef(Pbird.m)
# ranef(Pbird.m)

# mcmc_plot(Pbird.m, pars = "id_plant")
# launch_shinystan(Pvisit.m)

#DHARMa
Pbird.check <- check_brms(Pbird.m, integer = TRUE)
plot(Pbird.check, form = seeds.mh.sp$bird_sp)
plot(Pbird.check, form = seeds.mh.sp$mh)
testDispersion(Pbird.check)

saveRDS(Pbird.m, here("analysis", "output", paste0("Pbird.mh.model.rds")))
```



Predict:

```{r }
## Full posterior distributions

newdt <- expand.grid(nseed.total = 1, 
                     mh = sort(unique(seeds.mh.sp$mh)),
                     bird_sp = sort(unique(seeds.mh.sp$bird_sp))
)

Pbird.post <- posterior_linpred(Pbird.m, 
                                newdata = newdt,
                                ndraws = 1000) %>% 
  as.data.frame() %>%
  setNames(paste(newdt$mh, newdt$bird_sp, sep = "_")) %>% 
  tidyr::pivot_longer(cols = everything(), names_to = "mh.bird", values_to = "Pbird.logit") %>% 
  tidyr::separate(mh.bird, c("mh", "bird_sp"), sep = "_") %>% 
  arrange(mh, bird_sp) %>% 
  mutate(Pbird = plogis(Pbird.logit))


saveRDS(Pbird.post, here("analysis", "output", "Pbird.mh.post.rds"))

Pbird.post %>% 
  group_by(mh, bird_sp) %>% 
  summarise(mean.Pbird = mean(Pbird)) %>% 
  arrange(bird_sp, mh, desc(mean.Pbird)) %>% 
  kable(digits = 2)

```




## Calculate Number of seeds dispersed by each bird on each MH

The number of seeds from each bird species to each MH can be calculated as the product of the number of seeds arriving at each MH X Probability that seed came from a particular bird species (i.e. multiply posterior of model 1 and model 2)

```{r }
Nseed.post.long <- readRDS(here("analysis", "output", "Nseed.mh.post.rds"))
Pbird.post <- readRDS(here("analysis", "output", "Pbird.mh.post.rds"))
```


Calculate number of seeds arriving at each MH by each bird:

```{r}

n.mh.bird <- length(unique(seeds.mh.sp$mh))*length(unique(seeds.mh.sp$bird_sp)) 
# 5 mh * 27 bird spp = 135

Pbird.post <- Pbird.post %>% 
  mutate(seeds.m2 = case_when(
    mh == "PL" ~ rep(Nseed.post.long$seeds.m2[Nseed.post.long$mh == "PL"], times = n.mh.bird),
    mh == "FR" ~ rep(Nseed.post.long$seeds.m2[Nseed.post.long$mh == "FR"], times = n.mh.bird),
    mh == "NF" ~ rep(Nseed.post.long$seeds.m2[Nseed.post.long$mh == "NF"], times = n.mh.bird),
    mh == "OA" ~ rep(Nseed.post.long$seeds.m2[Nseed.post.long$mh == "OA"], times = n.mh.bird),
    mh == "PP" ~ rep(Nseed.post.long$seeds.m2[Nseed.post.long$mh == "PP"], times = n.mh.bird)
  )) %>% 
  mutate(seeds.m2.bird = seeds.m2 * Pbird)

Pbird.post %>% 
  group_by(bird_sp, mh) %>% 
  summarise(mean(seeds.m2.bird)) %>% 
  kable(digits = 1)

saveRDS(Pbird.post, here("analysis", "output", "seeds.m2.bird.mh.rds"))
```


Calculate, for each bird species, the Probability of dispersing seeds towards each MH:

```{r}
Pbird.list <- Pbird.post %>% 
  select(mh, bird_sp, seeds.m2.bird) %>% 
  split(f = ~bird_sp)

Pbird.mh.list <- lapply(Pbird.list, calc_prob_bird_mh)

Pbird.mh <- bind_rows(Pbird.mh.list)

saveRDS(Pbird.mh, here("analysis", "output", "Pdisperse.bird.mh.rds"))

Pbird.mh %>% 
  group_by(bird_sp, mh) %>% 
  summarise(median(Prob.bird.mh)) %>% 
  kable(digits = 2)
```




```{r}
sessioninfo::session_info()
```

