---
title: "SDE QUALITY"
date: "`r Sys.Date()`"
output: github_document
---


```{r setup, include=FALSE}
set.seed(888)
library(knitr)
knitr::opts_chunk$set(echo = TRUE, out.width = '100%', 
                      cache = TRUE, message = FALSE)
```

```{r }
library(here)
library(readr)
library(dplyr)
library(ggplot2)
suppressPackageStartupMessages(library(tidylog))

source(here("code", "functions.R"))

theme_set(theme_minimal())
```


----

Load components:

```{r}
seeds.mh <- readRDS(here("data", "Pdisperse.bird.mh.rds"))
pred.mh <- readRDS(here("data", "pred.mh.post.long.rds"))
surv.mh <- readRDS(here("data", "surv.mh.post.rds"))
```


Join data:

```{r}
n.mh.bird <- length(unique(seeds.mh$mh))*length(unique(seeds.mh$bird_sp)) 
# 5 mh * 27 bird spp = 135

qlty <- seeds.mh %>% 
  mutate(Prob.escape.rodent.pred = case_when(
    mh == "FR" ~ rep(pred.mh$Prob.escaping.pred[pred.mh$mh == "FR"], times = n.mh.bird),
    mh == "NF" ~ rep(pred.mh$Prob.escaping.pred[pred.mh$mh == "NF"], times = n.mh.bird),
    mh == "OA" ~ rep(pred.mh$Prob.escaping.pred[pred.mh$mh == "OA"], times = n.mh.bird),
    mh == "PL" ~ rep(pred.mh$Prob.escaping.pred[pred.mh$mh == "PL"], times = n.mh.bird),
    mh == "PP" ~ rep(pred.mh$Prob.escaping.pred[pred.mh$mh == "PP"], times = n.mh.bird)
  )) %>% 
  mutate(Prob.seedling.surv = case_when(
    mh == "FR" ~ rep(surv.mh$surv.avg[surv.mh$mh == "FR"], times = n.mh.bird),
    mh == "NF" ~ rep(surv.mh$surv.avg[surv.mh$mh == "NF"], times = n.mh.bird),
    mh == "OA" ~ rep(surv.mh$surv.avg[surv.mh$mh == "OA"], times = n.mh.bird),
    mh == "PL" ~ rep(surv.mh$surv.avg[surv.mh$mh == "PL"], times = n.mh.bird),
    mh == "PP" ~ rep(surv.mh$surv.avg[surv.mh$mh == "PP"], times = n.mh.bird)
  )) %>% 
  mutate(Prob.recruit.mh = Prob.escape.rodent.pred * Prob.seedling.surv) %>% 
  mutate(Prob.bird.recruit.mh = Prob.bird.mh * Prob.recruit.mh) %>% 
  select(bird_sp, mh, Prob.escape.rodent.pred, Prob.seedling.surv, Prob.recruit.mh,
         Prob.bird.disperse.mh = Prob.bird.mh, Prob.bird.recruit.mh)

```


```{r}
# Median
qlty %>% 
  group_by(mh) %>% 
  summarise(median(Prob.escape.rodent.pred), median(Prob.seedling.surv), median(Prob.recruit.mh))

# Mean
qlty %>% 
  group_by(mh) %>% 
  summarise(mean(Prob.escape.rodent.pred), mean(Prob.seedling.surv), mean(Prob.recruit.mh))

qlty %>% 
  group_by(mh) %>% 
  summarise(`escaping rodent predation` = mean(Prob.escape.rodent.pred), 
            `seedling survival` = mean(Prob.seedling.surv), 
            total = mean(Prob.recruit.mh)) %>% 
  pivot_longer(cols = -mh, names_to = "stage", values_to = "Probability") %>% 
  ggplot() +
  geom_line(aes(stage, Probability, group = mh, colour = mh), size = 2) 


ggplot(pred.mh) +
  geom_density(aes(Prob.escaping.pred), fill = "lightgrey") +
  facet_wrap(~mh, ncol = 1, scales = "free_y") +
  coord_cartesian(xlim = c(0, 0.05)) +
  theme(axis.text.y = element_blank()) +
  labs(title = "Probability of escaping rodent predation in each MH",
       x = "", y = "")

ggplot(surv.mh) +
  geom_density(aes(surv.avg), fill = "lightgrey") +
  facet_wrap(~mh, ncol = 1, scales = "free_y") +
  coord_cartesian(xlim = c(0, 0.1)) +
  theme(axis.text.y = element_blank()) +
  labs(title = "Probability of seedling surviving their 1st summer",
       x = "", y = "")
```


## Adjust SDE-QLTY for some species


### Keep only seeds escaping seed predators

For Chloris chloris, Pyrrhula pyrrhula, Fringilla coelebs and Coccothraustes that are seed predators, calculate how many of all the consumed seeds would escape predation (i.e. have any chance of producing a seedling).

Load the probability of escape:

```{r}
prob.escape <- readRDS(here("data", "prob.escape.rds"))
```

For all remaining bird species, the prob. of escape is 1

```{r}
qlty <- qlty %>% 
  group_by(bird_sp) %>% 
  mutate(Prob.escape.bird.pred = ifelse(
    bird_sp %in% c("Chloris chloris", "Pyrrhula pyrrhula", 
                   "Fringilla coelebs", "Coccothraustes coccothraustes"),
    rep(prob.escape$prob.escape, times = 5),  # 5 microhabitats 
    1)) %>% 
  ungroup() %>% 
  mutate(qlty.mh = Prob.escape.bird.pred * Prob.bird.recruit.mh)
```



*Columba palumbus* destroys all seeds in their gut.

*Cyanistes caeruleus* & *Parus major* are pulp predators. We assume here they don't drop/disperse any seed, so their SDE-QLTY is 0.


```{r}
qlty <- qlty %>% 
  mutate(qlty.mh = ifelse(bird_sp == "Columba palumbus", 0, qlty.mh),
         qlty.mh = ifelse(bird_sp == "Cyanistes caeruleus", 0, qlty.mh),
         qlty.mh = ifelse(bird_sp == "Parus major", 0, qlty.mh))
```




```{r}
# Do Prob. of dispersal to each MH adds up to 1 for each bird species?
qlty %>% 
 group_by(bird_sp) %>% 
 summarise(sum(Prob.bird.disperse.mh)) %>% 
  kable()


# Prob. dispersing to each MH per bird species
qlty %>% 
  group_by(bird_sp, mh) %>% 
  summarise(median(Prob.bird.disperse.mh)) %>% 
  kable(digits = 2)

# Prob., for each bird species, of delivering an entire seed to each MH and producing a seedling surviving their 1st summer. 
qlty %>% 
  group_by(bird_sp, mh) %>% 
  summarise(median(qlty.mh)) %>% 
  kable()



# Median:
qlty %>% 
  group_by(bird_sp) %>% 
  summarise(median(qlty.mh)) %>% 
  kable()

# Mean:
qlty %>% 
  group_by(bird_sp) %>% 
  summarise(mean(qlty.mh)) %>% 
  kable()
```




## Calculate overall probability of producing a recruit for each bird species (across all MH)

```{r}
sde.qlt <- qlty %>% 
  select(bird_sp, mh, qlty.mh) %>% 
  pivot_wider(names_from = "mh", values_from = "qlty.mh") %>% 
  tidyr::unnest(cols = c(FR, NF, OA, PL, PP)) %>% 
  mutate(P.recruit = FR + NF + OA + PL + PP)

sde.qlt %>% 
  group_by(bird_sp) %>% 
  summarise(median(P.recruit)) %>% 
  kable()
```


```{r}
sde.qlt %>% 
  filter(! bird_sp %in% c("Columba palumbus", "Cyanistes caeruleus", "Parus major")) %>% 
  ggplot() +
  geom_density(aes(P.recruit), fill = "lightgrey") +
  facet_wrap(~bird_sp, scales = "free_y") +
  coord_cartesian(xlim = c(0, 0.001)) +
  theme(axis.text.y = element_blank(),
        axis.text.x = element_text(size = 6)) +
  labs(x = "Probability of producing a recruit per dispersed seed") 


library(ggdist)

sde.qlt %>% 
  ggplot() +
  stat_pointinterval(aes(P.recruit, reorder(bird_sp, P.recruit, FUN = median))) +
  labs(title = "Seed disperser effectiveness (QUALITY component)",
       subtitle = "\nPosterior probability that a seed dispersed by each bird species\nproduces a seedling that survives their first summer\n",
       x = "\nProbability of producing a recruit",
       y = "",
       caption = "Points represent the median. Intervals represent 0.66 and 0.95 credibility intervals")
```


```{r}
saveRDS(qlty, here("analysis", "output", "SDE_QLTY_mh.rds"))
saveRDS(sde.qlt, here("analysis", "output", "SDE_QLTY_total.rds"))
```


```{r}
sessioninfo::session_info()
```

