---
title: "Calculate SDE QTY"
date: "`r Sys.Date()`"
output: github_document
---


```{r setup, include=FALSE}
set.seed(888)
library(knitr)
knitr::opts_chunk$set(echo = TRUE, out.width = '100%', 
                      cache = TRUE, message = FALSE)
```

```{r }
library(here)
library(dplyr)
library(tidyr)
library(ggplot2)
suppressPackageStartupMessages(library(tidylog))

source(here("code", "functions.R"))

theme_set(theme_minimal())
```




## SDE-QTY as Nfruits

Load data

```{r}
Nvisits.site <- readRDS(here("data", "Nvisits.site.rds"))

Pvisit.sp.Mad <- readRDS(here("data", "bc.Pvisit.sp.Madronas.rds")) %>%
  ungroup()
Pvisit.sp.Puntal <- readRDS(here("data", "PV.Puntal.combined.rds")) %>%
  ungroup() %>%
  mutate(site = "Puntal") %>%
  relocate(site)

Pfeed <- readRDS(here("data", "Pfeed.post.rds")) %>% 
  select(-Pfeed.logit) %>% 
  arrange(bird_sp)

Nfru <- readRDS(here("data", "Nfruits_visit_post_long.rds")) %>% 
  arrange(bird_sp)
```

Multiply:

```{r}
PV.sp.Mad <- Pvisit.sp.Mad %>% 
  mutate(Nvisit.site = rep(Nvisits.site$visits.Mad, times = nrow(Pvisit.sp.Mad)/1000)) %>% 
  select(site, id_plant, bird_sp, Pvisit.sp, Nvisit.site)

PV.sp.Puntal <- Pvisit.sp.Puntal %>% 
  mutate(Nvisit.site = rep(Nvisits.site$visits.Puntal, times = nrow(Pvisit.sp.Puntal)/1000)) %>% 
  select(site, id_plant, bird_sp, Pvisit.sp = PV.combined, Nvisit.site)

PV.sp <- bind_rows(PV.sp.Mad, PV.sp.Puntal)

PV.sp.list <- split(PV.sp, PV.sp$bird_sp)

PV.sp.etc <- lapply(PV.sp.list, add_Pfeed_Nfru_sp, pfeed = Pfeed, nfru = Nfru)

PV.sp.all <- bind_rows(PV.sp.etc) 

PV.sp.all <- PV.sp.all %>% 
  mutate(Nvisit.sp = Nvisit.site*Pvisit.sp) %>% 
  mutate(qty.fru = Nvisit.site*Pvisit.sp*Pfeed*Nfruits)

saveRDS(PV.sp.all, here("analysis", "output", "SDE_QTY_Nfruits.rds"))

# PV.sp.all %>% 
#   group_by(site, id_plant, bird_sp) %>% 
#   summarise(mean.sde.qty = round(mean(qty.fru), digits = 3)) %>% 
#   arrange(desc(mean.sde.qty)) %>% 
#   kable()
```



Plot number of visits:

```{r}
library(ggdist)

PV.sp.all %>% 
  ggplot() +
  facet_wrap(~id_plant, ncol = 5) +
  aes(x = reorder(bird_sp, Nvisit.sp), y = Nvisit.sp) +
  stat_pointinterval(size = 1, .width = 0.50) +
  coord_flip(ylim = c(0, 7000)) +
  theme(axis.text.y = element_text(size = 5),
        axis.text.x = element_text(size = 10)) +
  labs(title = "Number of visits by each bird species to each plant",
       subtitle = "2019-2020",
       x = "",
       y = "Visits",
       caption = "Note axis has been truncated at 7000") 

ggsave(here("analysis", "output", "Nvisits.sp_final.pdf"), width = 15, height = 40, units = "in")

```



Plot number of fruits consumed per bird species

```{r}
ggplot(Nfru) + 
  facet_wrap(~bird_sp, ncol = 4, scales = "free") +
  geom_histogram(aes(Nfruits), binwidth = 1) +
  theme(axis.text.y = element_blank()) +
  labs(x = "", y = "",
       title = "Number of fruits consumed per visit")

ggsave(here("analysis", "output", "Nfruits_visit.pdf"), width = 15, height = 30, units = "in")
  
```



Plot total number of fruits consumed by each bird species from each plant:

```{r}
library(ggdist)

PV.sp.all %>% 
  ggplot() +
  facet_wrap(~id_plant, ncol = 5) +
  aes(x = reorder(bird_sp, qty.fru), y = qty.fru) +
  stat_pointinterval(size = 1, .width = 0.50) +
  coord_flip(ylim = c(0, 10000)) +
  theme(axis.text.y = element_text(size = 5),
        axis.text.x = element_text(size = 10)) +
  labs(title = "Number of fruits consumed by each bird species from each plant",
       subtitle = "2019-2020",
       x = "",
       y = "",
       caption = "Note axis has been truncated at 10000") 

ggsave(here("analysis", "output", "SDE_QTY_Nfruits.pdf"), width = 15, height = 40, units = "in")



## Zoom 
PV.sp.all %>% 
  ggplot() +
  facet_wrap(~id_plant, ncol = 5) +
  aes(x = reorder(bird_sp, qty.fru), y = qty.fru) +
  stat_pointinterval(size = 1, .width = 0.50) +
  coord_flip(ylim = c(0, 1000)) +
  theme(axis.text.y = element_text(size = 5),
        axis.text.x = element_text(size = 10)) +
  labs(title = "Number of fruits consumed by each bird species from each plant",
       # subtitle = "",
       x = "",
       y = "",
       caption = "Note axis has been truncated at 1000") 

ggsave(here("analysis", "output", "SDE_QTY_Nfruits_zoom.pdf"), width = 15, height = 40, units = "in")
```



## Keep only seeds escaping seed predators

For *Chloris chloris*, *Pyrrhula pyrrhula*, *Fringilla coelebs* and *Coccothraustes* that are seed predators, calculate how many of all the consumed seeds would escape predation (i.e. have any chance of producing a seedling).

Load the probability of escape:

```{r}
prob.escape <- readRDS(here("data", "prob.escape.rds"))
```

For all remaining bird species, the prob. of escape is 1

```{r}
PV.sp.all <- PV.sp.all %>% 
  group_by(site, id_plant, bird_sp) %>% 
  mutate(prob.escape = ifelse(
    bird_sp %in% c("Chloris chloris", "Pyrrhula pyrrhula", 
                   "Fringilla coelebs", "Coccothraustes coccothraustes"),
    prob.escape$prob.escape, 
    1)) %>% 
  ungroup() %>% 
  mutate(qty.fru.escape = qty.fru * prob.escape)

saveRDS(PV.sp.all, here("analysis", "output", "SDE_QTY_Nfruits_escape.rds"))
```



---------------

```{r}
sessioninfo::session_info()
```

