---
title: "Contribution of Quantity and Quality components to the variation of effectiveness"
output: github_document
---

```{r, warning=F, message=F}
library(here)
library(tidyverse)
```

Read data (n=1000 values per interaction, i.e., posteriors):
```{r}
eff <- readRDS(here("data/effectiveness.rds")) %>%
  mutate(qty.fru.manipulated=as.numeric(qty.fru.manipulated)) %>% filter(!is.na(fruit.reward))
```

Remove non-legitimate dispersers for calculating variance in SDE
```{r}
non.disp <- c("Chloris chloris", "Cyanistes caeruleus", "Fringilla coelebs","Parus major",
                    "Pyrrhula pyrrhula","Coccothraustes coccothraustes", "Columba palumbus")

eff.sd <- eff %>% filter(!bird_sp %in% non.disp) 
```

### Component distribution:
```{r, include=F}
summary(eff$qty.fru.consumed) #RPE-QTC
summary(eff$fruit.reward) #RPE-QLC
summary(eff$qty.fru.manipulated) #SDE-QTC
summary(eff$sde.qlc) #SDE-QLC
```


### Linear models and variance partitioning

**For RPE:**

Full model, where R-squared is 1
```{r}
m1 <- lm(log(rpe) ~ log(qty.fru.consumed) + log(fruit.reward), data=eff)
summary(m1)
```

RPE~QTC model
```{r}
m2 <- lm(log(rpe) ~ log(qty.fru.consumed), data=eff)
m2.sum <- summary(m2)
m2.sum
```

RPE~QLC model
```{r}
m3 <- lm(log(rpe) ~ log(fruit.reward), data=eff)
m3.sum <- summary(m3)
m3.sum
```

Normalise values to add up to 100%:
```{r}
#Quantity
m2.sum$r.squared / sum(m2.sum$r.squared, m3.sum$r.squared)

#Quality
m3.sum$r.squared / sum(m2.sum$r.squared, m3.sum$r.squared)
```


**For SDE:**

Full model, where R-squared is 1
```{r}
m4 <- lm(log(sde) ~ log(qty.fru.manipulated) + log(sde.qlc), data=eff.sd)
summary(m4)
```

SDE~QTC model
```{r}
m5 <- lm(log(sde) ~ log(qty.fru.manipulated), data=eff.sd)
m5.sum <- summary(m5)
m5.sum
```

SDE~QLC model
```{r}
m6 <- lm(log(sde) ~ log(sde.qlc), data=eff.sd)
m6.sum <- summary(m6)
m6.sum
```

Normalise values to add up to 100%:
```{r}
#Quantity
m5.sum$r.squared / sum(m5.sum$r.squared, m6.sum$r.squared)

#Quality
m6.sum$r.squared / sum(m5.sum$r.squared, m6.sum$r.squared)
```

