---
title: "Calculating Probability of bird visit from videos (only for Puntal)"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
set.seed(888)
library(knitr)
knitr::opts_chunk$set(echo = TRUE, out.width = '100%', cache = TRUE, message = FALSE)
```

```{r SDE-QTY-PV-Videos-1 }
library(here)
library(tidyr)
library(dplyr)
library(ggplot2)
library(brms)
library(DHARMa)
library(ggdist)

source(here("code", "functions.R"))

theme_set(theme_minimal())
```

Load data:

```{r SDE-QTY-PV-Videos-2 }
camdata <- readr::read_csv(here("data", "cam_indiv_lentisc.csv"))
camdata
glimpse(camdata)
camdata$site <- "Puntal"
```


```{r SDE-QTY-PV-Videos-3, results = "asis"}
summarytools::dfSummary(camdata, plain.ascii = FALSE, style = "grid", valid.col = FALSE)
```


## Calculating the QUANTITY component of Seed Dispersal Effectiveness (SDE)


### 1. Nvisit

Here we estimate the number of bird visits to each individual plant (per hour of recording) over the whole fruiting season. 

So far we count all visits without considering if there was consumption or not.


#### 1.1 Prepare data for model & EDA

Number of recording hours per individual:

```{r SDE-QTY-PV-Videos-4}
rectime.indiv <- camdata %>% 
  group_by(id_plant) %>% 
  distinct(date_recording, .keep_all = TRUE) %>% 
  summarise(rec_time_indiv = sum(rec_time_hours)) 

ggplot(rectime.indiv) +
  geom_col(aes(reorder(id_plant, rec_time_indiv), rec_time_indiv)) +
  coord_flip() +
  labs(x = "Individual", y = "Total recording hours")

camdata <- camdata %>% 
  left_join(rectime.indiv)
  
```


Recorded visits per individual lentisc:

```{r SDE-QTY-PV-Videos-5, fig.height=5}
camdata.visits <- camdata %>% 
  filter(!is.na(bird_sp)) %>%  # NA means no visit at all in that recording session
  group_by(id_plant, rec_time_indiv) %>% 
  summarise(visits.ind = n()) %>%  # including 'unknown' = unidentified bird
  arrange(desc(visits.ind))
  

ggplot(camdata.visits) +
  geom_col(aes(reorder(id_plant, visits.ind), visits.ind)) +
  coord_flip() +
  labs(x = "Individual", y = "Number of visits", 
       title = "Number of recorded bird visits over the 2019-20 season",
       subtitle = "Puntal site")
```


Number of visits per bird species throughout the period:

```{r SDE-QTY-PV-Videos-6, fig.height=6}
camdata %>% 
  group_by(bird_sp, cam_period) %>% 
  summarise(nvisit = n()) %>% 
  ggplot() +
  facet_wrap(~bird_sp) + 
  geom_col(aes(cam_period, nvisit)) +
  theme(strip.text = element_text(size = 8))
```




#### 1.2 Model

Using Poisson to model Nvisits per plant per hour of recording:

```{r Nvisit-formu}

# rate(rec_time_indiv) = offset(log(rec_time_indiv)), see ?bf
Nvisit.formu <- brmsformula(visits.ind | rate(rec_time_indiv) ~  
                             (1 | id_plant), 
                            family = poisson())

get_prior(Nvisit.formu, data = camdata.visits)

## priors (log scale)
# exp(1.4) = 4.05 visits/h
# plot(density(rlnorm(1000, 1.4, 1)))
priors <- c(
  set_prior("normal(1.4, 1)", class = "Intercept"), 
  set_prior("normal(0, 1)", class = "sd")
)

```


Prior predictive check:

```{r Nvisit-ppc}

Nvisit.m <- brm(Nvisit.formu,
         data = camdata.visits,
         prior = priors,
         sample_prior = "only",
         cores = 4,
         iter = 2000,
         refresh = 0)

# stancode(Nvisit.m)
pp_check(Nvisit.m, nsamples = 100) + coord_cartesian(xlim = c(0, 500))
summary(Nvisit.m, priors = TRUE)
plot(Nvisit.m)
```



Fit model:

```{r Nvisit-fit}
Nvisit.m <- brm(Nvisit.formu,
         data = camdata.visits,
         prior = priors,
         # sample_prior = "only",
         cores = 4,
         iter = 4000,
         refresh = 0)

```


```{r Nvisit-check}

pp_check(Nvisit.m, nsamples = 100)
bayesplot::ppc_rootogram(y = camdata.visits$visits.ind, 
                         yrep = posterior_predict(Nvisit.m, nsamples = 1000))
# pp_check(Nvisit.m, nsamples = 100, type = "error_scatter_avg")
bayestestR::check_prior(Nvisit.m)

summary(Nvisit.m, priors = TRUE)

plot(Nvisit.m)

stancode(Nvisit.m)

# coef(Nvisit.m)
fixef(Nvisit.m)
lapply(ranef(Nvisit.m), round, digits = 3)

#mcmc_plot(Nvisit.m, pars = "id_plant")
# launch_shinystan(Nvisit.m)

#DHARMa
Nvisit.check <- check_brms(Nvisit.m, integer = TRUE)
testDispersion(Nvisit.check)

saveRDS(Nvisit.m, here("analysis", "output", "cam.Nvisit.model.rds"))

```



Predict:

```{r SDE-QTY-PV-Videos-11 }
## Full posterior distributions
cam.visits.post <- posterior_predict(Nvisit.m, 
                                    newdata = data.frame(
                                    id_plant = camdata.visits$id_plant,
                                    rec_time_indiv = 1),
                                    nsamples = 1000) %>% 
  as.data.frame() %>%
  setNames(unique(camdata.visits$id_plant))

cam.visits.post.long <- cam.visits.post %>% 
  tidyr::pivot_longer(cols = everything(), names_to = "id_plant", values_to = "visits") %>% 
  mutate(id_plant = as.numeric(id_plant)) %>% 
  right_join(select(camdata.visits, id_plant), by = "id_plant") %>% 
  arrange(id_plant) %>% 
  mutate(id_plant = as.character(id_plant))

saveRDS(cam.visits.post.long, here("analysis", "output", "cam.indiv.Puntal.Nvisits.post.long.rds"))

```


Sum total number of fruits for each iteration to calculate prop of visit to each individual:

```{r }
cam.Pvisit.post <- Nvisit2Pvisit(cam.visits.post, tidy = TRUE) %>% 
  left_join(select(camdata.visits, id_plant), by = "id_plant") %>% 
  arrange(id_plant)

cam.Pvisit.post %>% 
  group_by(id_plant) %>% 
  summarise(mean(Pvisit)) %>% 
  arrange(desc(`mean(Pvisit)`)) %>% 
  kable(digits = 2)

saveRDS(cam.Pvisit.post, here("analysis", "output", 
                                paste0("cam.indiv.Puntal.Pvisit.post.long.rds")))

```



Plot:

```{r SDE-QTY-PV-Videos-12 }

cam.visits.post.long %>% 
  group_by(id_plant) %>% 
  ggplot() +
  aes(x = reorder(id_plant, visits), y = visits) +
  stat_gradientinterval() +
  coord_flip() +
  labs(x = "Plant", y = "Visits/h")

ggsave(here("analysis", "output", "cam.visits.post.pdf"), 
       device = "pdf", width = 9, height = 11, units = "in")
```



------------------------------------------------------------------

### 2. Pbird: Probability of visit by each bird species at each plant

#### 2.1 Prepare data

Get number of visits per bird species per plant. 

We remove the unknown (unidentified birds) for both the numerator and denominator as they could be any bird species. 


```{r SDE-QTY-PV-Videos-13 }
camdata.visits.sp <- camdata %>% 
  filter(!is.na(bird_sp)) %>%  # Video without any visit observed
  filter(bird_sp != "unknown") %>% 
  group_by(id_plant, rec_time_indiv, bird_sp) %>% 
  summarise(visits.bird.plant = n()) %>% 
  ungroup() %>% 
  mutate(site = "Puntal_cam") %>% 
  relocate(site)
```


Add interactions that were observed in barcoding but not in videos (only in Puntal):

```{r Barcoding-PV-indiv-bird-19}
bcdata <- readr::read_csv(here("data", "bc_indiv_lentisc.csv")) %>% 
  filter(site == "Puntal")

bc.only <- bcdata %>% 
  select(id_plant, bird_sp) %>% 
  filter(!is.na(bird_sp), bird_sp != "unknown") %>% 
  distinct() %>% 
  anti_join(select(camdata.visits.sp, id_plant, bird_sp)) %>% 
  mutate(site = "Puntal_bc", 
         visits.bird.plant = 0) %>% 
  left_join(distinct(select(camdata.visits.sp, id_plant, rec_time_indiv)), by = "id_plant")

kable(select(bc.only, id_plant, bird_sp))

camdata.visits.sp <- full_join(camdata.visits.sp, bc.only)
```



Add other unobserved interactions (value = 0):

```{r}
all.int <- expand.grid(id_plant = sort(unique(camdata.visits.sp$id_plant)), 
                       bird_sp = sort(unique(camdata.visits.sp$bird_sp)), 
                       stringsAsFactors = FALSE)

unobs <- anti_join(all.int, select(camdata.visits.sp, id_plant, bird_sp)) %>% 
    mutate(site = "Puntal_unobs", 
           visits.bird.plant = 0) %>% 
  relocate(site) %>% 
  left_join(distinct(select(camdata.visits.sp, id_plant, rec_time_indiv)), by = "id_plant") %>% 
  relocate(rec_time_indiv, .before = bird_sp)

camdata.visits.sp <- full_join(camdata.visits.sp, unobs)
```





Calculate total number of **identified** visits per plant: 

```{r}

camdata.visits.id <- camdata.visits.sp %>% 
  group_by(id_plant) %>% 
  summarise(visits.plant.total = sum(visits.bird.plant)) %>%  
  arrange(desc(visits.plant.total))

camdata.visits.sp <- camdata.visits.sp %>% 
  left_join(camdata.visits.id, by = "id_plant") %>% 
  mutate(visits.bird.plant.prop = visits.bird.plant/visits.plant.total) %>% 
  ungroup() %>% 
  arrange(id_plant, bird_sp)

camdata.visits.sp$row <- 1:nrow(camdata.visits.sp)

```


Number of visits per bird species: 

```{r SDE-QTY-PV-Videos-14 }
camdata.visits.sp %>% 
  group_by(bird_sp) %>% 
  summarise(suma = sum(visits.bird.plant)) %>% 
  arrange(desc(suma)) %>% 
  kable()
```




#### 2.2 Model

Prior predictive check:

```{r Pbird-formu}

Pbird.formu <- bf(visits.bird.plant | trials(visits.plant.total) ~
                    (1 | bird_sp) + (1 | id_plant) + 
                    (1 | bird_sp:id_plant),
                  family = binomial())

get_prior(Pbird.formu, data = camdata.visits.sp)

# priors (logit scale)

# 1/26 bird species = 0.038

# plogis(-2) = 0.12   
# plogis(-3) = 0.047
# plogis(-1) = 0.27
# plogis(-4) = 0.018
# plogis(0) = 0.5
priors <- c(
  set_prior("normal(-6.5, 1)", class = "Intercept"), 
  set_prior("normal(0, 3)", class = "sd")
)
```

Prior predictive check:

```{r Pbird-ppc}
Pbird.m <- brm(Pbird.formu,
         data = camdata.visits.sp,
         prior = priors,
         sample_prior = "only",
         cores = 4,
         iter = 2000,
         refresh = 0)

# stancode(Nvisit.m)
# pp_check(Pbird.m, nsamples = 100)
pp_check(Pbird.m, nsamples = 100) + coord_cartesian(xlim = c(0, 5))
summary(Pbird.m, priors = TRUE)
plot(Pbird.m)

```


Fit model:

```{r Pbird-fit}
Pbird.m <- brm(Pbird.formu,
         data = camdata.visits.sp,
         prior = priors,
         cores = 4,
         iter = 4000,
         refresh = 0)
```


Check:

```{r Pbird-check}
pp_check(Pbird.m, nsamples = 100)
pp_check(Pbird.m, nsamples = 100) + coord_cartesian(xlim = c(0, 5))
pp_check(Pbird.m, nsamples = 100, type = "bars") + coord_cartesian(xlim = c(0, 5))
# pp_check(Pvisit.m, nsamples = 100, type = "error_scatter_avg")
suppressWarnings(bayestestR::check_prior(Pbird.m))

summary(Pbird.m, priors = TRUE)

plot(Pbird.m)

stancode(Pbird.m)

lapply(coef(Pbird.m), head)
fixef(Pbird.m)
# ranef(Pbird.m)

# mcmc_plot(Pbird.m, pars = "id_plant")
# launch_shinystan(Pvisit.m)

#DHARMa
Pbird.check <- check_brms(Pbird.m, integer = TRUE)
plot(Pbird.check, form = camdata.visits.sp$bird_sp)
plot(Pbird.check, form = camdata.visits.sp$id_plant)
testDispersion(Pbird.check)

saveRDS(Pbird.m, here("analysis", "output", "cam.Pbird.model.rds"))
```



Predict:

```{r SDE-QTY-PV-Videos-17 }

## Full posterior distributions
only.obs <- filter(camdata.visits.sp, site != "Puntal_unobs")

cam.Pbird.post <- posterior_linpred(Pbird.m, 
                                    newdata = only.obs,
                                    nsamples = 1000) %>% 
  as.data.frame() %>%
  setNames(paste(only.obs$id_plant, only.obs$bird_sp, sep = "_")) %>% 
  tidyr::pivot_longer(cols = everything(), names_to = "plant.bird", values_to = "Pbird.logit") %>% 
  tidyr::separate(plant.bird, c("id_plant", "bird_sp"), sep = "_") %>% 
  mutate(id_plant = as.numeric(id_plant)) %>% 
  left_join(distinct(select(only.obs, id_plant)), by = "id_plant") %>% 
  arrange(id_plant, bird_sp) %>% 
  mutate(id_plant = as.character(id_plant)) %>% 
  mutate(Pbird = plogis(Pbird.logit))

saveRDS(cam.Pbird.post, here("analysis", "output", "cam.Pbird.post.rds"))

cam.Pbird.post %>% 
  group_by(id_plant, bird_sp) %>% 
  summarise(mean.Pbird = mean(Pbird)) %>% 
  arrange(id_plant, desc(mean.Pbird)) %>% 
  kable(digits = 2)

```




# Plots


```{r SDE-QTY-PV-Videos-20, fig.height=6}

plot_Pbird_plant(pop = "Puntal", df = cam.Pbird.post, bc.cam = "cam")

plot_Pbird_bird(pop = "Puntal", df = cam.Pbird.post, bc.cam = "cam")

```



### 3. Calculate Pvisit for each bird at each plant

The number of visits from each bird species to each individual plant can be calculated as the product of the total visits to each plant X Probability that a visit is by a particular species (i.e. multiply posterior of model 1 and model 2)

```{r SDE-QTY-PV-Videos-21 }
# Nvisit <- readRDS(here("analysis", "output", "cam.visits.post.rds"))
Pvisit <- readRDS(here("analysis", "output", "cam.indiv.Puntal.Pvisit.post.long.rds"))
Pbird <- readRDS(here("analysis", "output", "cam.Pbird.post.rds"))
```


```{r SDE-QTY-PV-Videos-23 }

Pbird.list <- split(select(Pbird, -Pbird.logit), ~id_plant)

Pvisit.sp.list <- lapply(Pbird.list, mutiply_PVindiv_Pbird, Pvis = Pvisit)

Pvisit.sp <- bind_rows(Pvisit.sp.list) %>% 
  rename(PVindiv = Pv) %>% 
  relocate(PVindiv, .before = Pbird)

saveRDS(Pvisit.sp, here("analysis", "output", "cam.Pvisit.sp.Puntal.rds"))
```



```{r SDE-QTY-PV-Videos-25, fig.height=7}
plot_PV_plant(pop = "Puntal", df = Pvisit.sp, bc.cam = "cam")
```


-------------------------------------------------

```{r}
sessioninfo::session_info()
```


